### 2015-03-09 article/abstract.tex
article initial commit

Nuclear receptors (NRs) are important signaling mediators in cells due to their ability to act as receptors and transcription factors. These molecules are activated by ligands and regulate critical functions in cell control, inflammation, fibrosis and tumor formation. Nuclear receptors act on the metabolism and signaling of cells by changing the expression of target genes. Here, we will investigate the knockout phenotypes and genetic variation of mouse NRs. We will first create an assembly of all known mouse SNPs in the vicinity of mouse NR genes (+-1Mbp up-or downstream). In a second step, phenotype information for genetic knockouts and genetic variation data will be included. Knockout phenotypes are available from the MGI database, while the Mouse Phenome Database provides SNPs from various mouse strains, which can be correlated to extreme phenotypes, measured in these mouse strains. The goal of this analysis is to find NR SNPs in mice that influence changes in biological parameters such as body weight, body fat and other phenotypic traits. Furthermore, we will couple these findings to phenotypes observed in mice with a targeted or spontaneous mutation of the nuclear receptor and thus provide additional indication for a putative functionality of the investigated SNPs.

### 2015-03-09 article/article.tex
article initial commit

\documentclass[fleqn,11pt]{SelfArx} 
%\setlength{\fboxrule}{0.75pt} % Width of the border around the abstract
\definecolor{color1}{RGB}{0,0,100} % Color of the article title and sections
\definecolor{color2}{RGB}{0,0,0} % Color of the boxes behind the abstract and headings
\usepackage{hyperref} % Required for hyperlinks
\hypersetup{hidelinks,colorlinks,breaklinks=true,urlcolor=color2,citecolor=color1,linkcolor=color1,bookmarksopen=false,pdftitle={Title},pdfauthor={Author}}

%----------------------------------------------------------------------------------------
%	ARTICLE INFORMATION
%----------------------------------------------------------------------------------------
\PaperTitle{Nuclear receptor variation in mice} % Article title

\Authors{Beratis, Alexander, Iacob, Diana, Boyanova, Dr. Desislava\textsuperscript{1}, Mewes, Prof. Dr. Hans-Werner\textsuperscript{1}} % Authors
\affiliation{\textsuperscript{1}\textit{Institute of Bioinformatics and Systems Biology, Helmholtz Zentrum MŸnchen, German Research Center for Environmental Health,}} % Author affiliation

\Keywords{Nuclear receptors --- SNPs --- Gene variation} % Keywords
\newcommand{\keywordname}{Keywords} % Defines the keywords heading name

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\Abstract{\textit{Nuclear receptors (NRs) are important signaling mediators in cells due to their ability to act as receptors and transcription factors. These molecules are activated by ligands and regulate critical functions in cell control, inflammation, fibrosis and tumor formation. Nuclear receptors act on the metabolism and signaling of cells by changing the expression of target genes. Here, we will investigate the knockout phenotypes and genetic variation of mouse NRs. We will first create an assembly of all known mouse SNPs in the vicinity of mouse NR genes (+-1Mbp up-or downstream). In a second step, phenotype information for genetic knockouts and genetic variation data will be included. Knockout phenotypes are available from the MGI database, while the Mouse Phenome Database provides SNPs from various mouse strains, which can be correlated to extreme phenotypes, measured in these mouse strains. The goal of this analysis is to find NR SNPs in mice that influence changes in biological parameters such as body weight, body fat and other phenotypic traits. Furthermore, we will couple these findings to phenotypes observed in mice with a targeted or spontaneous mutation of the nuclear receptor and thus provide additional indication for a putative functionality of the investigated SNPs.}}

%----------------------------------------------------------------------------------------

\begin{document}

\flushbottom % Makes all text pages the same height
\maketitle % Print the title and abstract box
\thispagestyle{empty} % Removes page numbering from the first page

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section*{Introduction} % The \section*{} command stops section numbering

IntroIntroIntro

%------------------------------------------------

\section{Methods}

\subsection{Nuclear receptors}
49 nuclear receptors
\subsection{human genome}
tbd

%------------------------------------------------

\section{Tools}
\subsection{MGI}
Mouse Genome Informatics\footnote{\url{http://www.informatics.jax.org/}, \today}.
\subsection{MPD}
Mouse Phenome Database\footnote{\url{http://phenome.jax.org/}, \today}~\cite{mpd}.

%------------------------------------------------

\section{Database}

%------------------------------------------------

\section{Results}

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{pics/mgi_phenotypes_distribution.pdf}
	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
 	\caption{\footnotesize{\textbf{MGI phenotypes}. 
	~~~~~~~\\
	Mouse Genome Informatics phenotype distribution over the nuclear receptors.}}
	\label{fig:mgi_pheotypes_distribution}
\end{figure}
~~~~~~~\\

%Figure~\ref{fig:mgi_pheotypes_distribution} bla bla 
\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{pics/mgi_types_distribution.pdf}
	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
 	\caption{\footnotesize{\textbf{MGI allele types}. 
	~~~~~~~\\
	Mouse Genome Informatics allele type distribution over the nuclear receptors.}}
	\label{fig:mgi_types_distribution}
\end{figure}
~~~~~~~\\

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{pics/mpi_phenotypes_distribution_zscore2.pdf}
	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
 	\caption{\footnotesize{\textbf{MPD phenotypes, having Z-score $>$ 2}. 
	~~~~~~~\\
	Mouse Phenotype Database phenotype distribution over the nuclear receptors, having Z-score $>$ 2}}
	\label{fig:mpi_pheotypes_distribution_zscore2}
\end{figure}
~~~~~~~\\

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{pics/mpi_phenotypes_distribution_zscore2_neg.pdf}
	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
 	\caption{\footnotesize{\textbf{MPD phenotypes, having Z-score $<$ -2}. 
	~~~~~~~\\
	Mouse Phenotype Database phenotype distribution over the nuclear receptors, having Z-score $<$ -2}}
	\label{fig:mpi_pheotypes_distribution_zscore2_neg}
\end{figure}
~~~~~~~\\

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{pics/mgi_phenotypes_nr.pdf}
	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
 	\caption{\footnotesize{\textbf{MGI phenotype - nuclear receptor associations}. 
	~~~~~~~\\
	Mouse Genome Informatics phenotype occurrence frequency among the nuclear receptors.}}
	\label{fig:mgi_pheotypes_nr}
\end{figure}
~~~~~~~\\

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{pics/mgi_types_nr.pdf}
	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
 	\caption{\footnotesize{\textbf{MGI allele type - nuclear receptor associations}. 
	~~~~~~~\\
	Mouse Genome Informatics allele type occurrence frequency among the nuclear receptors.}}
	\label{fig:mgi_types_nr}
\end{figure}
~~~~~~~\\

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{pics/mpi_phenotypes_strain_zscore2.pdf}
	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
 	\caption{\footnotesize{\textbf{MPD phenotype - strain associations, having Z-score $>$ 2}. 
	~~~~~~~\\
	Mouse Phenome Database phenotype occurrence frequency among the strains associated with the nuclear receptors, having Z-score $>$ 2}}
	\label{fig:mgi_pheotypes_strain_zscore2}
\end{figure}
~~~~~~~\\

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{pics/mpi_phenotypes_strain_zscore2.pdf}
	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
 	\caption{\footnotesize{\textbf{MPD phenotype - strain associations, having Z-score $<$ -2}. 
	~~~~~~~\\
	Mouse Phenome Database phenotype occurrence frequency among the strains associated with the nuclear receptors, having Z-score $<$ -2}}
	\label{fig:mgi_pheotypes_strain_zscore2_neg}
\end{figure}

%------------------------------------------------
\section{Discussion and Outlook}

Blabla

%------------------------------------------------
\section{Acknowledgments} % The \section*{} command stops section numbering

Our gratitude to Prof. Dr. H. W. Mewes for offering us the topic and the opportunity to work with the Helmholz Zentrum research centre in Munich. Many thanks to our supervisor, Dr. Desislava Boyanova for all the input and ideas, discussions, advices and foremost her useful and critical suggestions which motivated us a lot. Last but not least, we would like to thank our fellow colleagues for their support, as well as the entire \emph{Helmholz Zentrum} group. 
%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\phantomsection
\bibliographystyle{unsrt}
\bibliography{sample}

%----------------------------------------------------------------------------------------

\end{document}

### 2015-03-09 plotPipeline.pl

@@ -24,7 +24,7 @@ pod2usage( -verbose => 2 ) if $man;
 usage() unless defined $outputDir;
 
 make_path ($outputDir);
-
+=co
 my $mgi_phenotypes_distribution_file = $outputDir."/mgi_phenotypes_distribution.pdf";
 system (" perl ./analysisMGI_phenotypes_distribution.pl --outputFile $mgi_phenotypes_distribution_file");
 
@@ -37,13 +37,12 @@ system (" perl ./analysisMPI_phenotypes_distribution_zscore2.pl --outputFile $mp
 my $mpi_phenotypes_distribution_zscore2_neg_file = $outputDir."/mpi_phenotypes_distribution_zscore2_neg.pdf";
  system (" perl ./analysisMPI_phenotypes_distribution_zscore-2.pl --outputFile $mpi_phenotypes_distribution_zscore2_neg_file");
 
-my $mpi_snp_distribution_file = $outputDir."/mpi_snp_distribution.pdf";
-system (" perl ./analysisMPI_snp_distribution.pl --outputFile $mpi_snp_distribution_file");
+#my $mpi_snp_distribution_file = $outputDir."/mpi_snp_distribution.pdf";
+#system (" perl ./analysisMPI_snp_distribution.pl --outputFile $mpi_snp_distribution_file");
 
 my $mpi_snp_function_distribution_file = $outputDir."/mpi_snp_function_distribution.pdf";
 system (" perl ./analysisMPI_snp_function_distribution.pl --outputFile $mpi_snp_function_distribution_file");
-
-=co
+=cut
 my $mgi_phenotypes_nr_file = $outputDir."/mgi_phenotypes_nr.pdf";
 system (" perl ./analysisMGI_phenotypes_nr.pl --outputFile $mgi_phenotypes_nr_file");
 
@@ -56,6 +55,7 @@ system (" perl ./analysisMPI_phenotypes_strain_zscore2.pl --outputFile $mpi_phen
 my $mpi_phenotypes_strain_zscore2_neg_file = $outputDir."/mpi_phenotypes_strain_zscore2_neg.pdf";
 system (" perl ./analysisMPI_phenotypes_strain_zscore-2.pl --outputFile $mpi_phenotypes_strain_zscore2_neg_file");
 
+=co
 my $mpi_phenotypes_nr_zscore2_file = $outputDir."/mpi_phenotypes_nr_zscore2.pdf";
[…]

### 2015-03-09 article/article.tex
commit

@@
 \subsection{MGI}
-Mouse Genome Informatics\footnote{\url{http://www.informatics.jax.org/}, \today}.
+Mouse Genome Informatics\footnote{\url{http://www.informatics.jax.org/}, \today} is a database for the laboratory mouse, which makes information about integrated genetics and associated phenotypes and alleles available.
+The 49 nuclear receptors were searched in the MGI database to achieve an association between these nuclear receptors and phenotypes which arise from alternation in the scope of their genes. 
 \subsection{MPD}

### 2015-03-09 article/article.tex
commit

--- 2015-03-09 article/article.tex
@@
 \subsection{MGI}
-Mouse Genome Informatics\footnote{\url{http://www.informatics.jax.org/}, \today}.
+Mouse Genome Informatics\footnote{\url{http://www.informatics.jax.org/}, \today} is a database for the laboratory mouse, which makes information about integrated genetics and associated phenotypes with their alleles available.
+For this report the MGI database was solely used for building a connection between the nuclear receptor genes and the associated phenotypes always dependent on miscellaneous strains.
+Information about SNPs and variation of the nuclear receptors were not taken from MGI, because the annotation regarding this matter was far better in the MPI and UCSC database.
 \subsection{MPD}
-Mouse Phenome Database\footnote{\url{http://phenome.jax.org/}, \today}~\cite{mpd}.
+Mouse Phenome Database\footnote{\url{http://phenome.jax.org/}, \today}~\cite{mpd} includes information about measured data on laboratoy mouse strains and populations as well as SNPs and phenotypes of the examined strains.
+The information about the SNPs and the variations were the motifs to also take a look at this database, wit
+Given that the MGI phenotypes can be compared to the ones found in MPI which are by far more detailed.
+The drawback of the MPI database is that every sort of data is always reffered on the regarding strain, making the mapping between the nuclear receptors and the MGI data more difficult.
 
@@
 \section{Database}
+The goal of the database was to create several tables which could be mapped with the reference table of the 49 nuclear receptors, to create statistics about the relationship between the nuclear receptors and the phenotypes and SNPs associated with them.
+As mentioned above, the main problem was the connection between the MGI and the MPI tables, which was gone around	by using an extra table which contained only the strains of MGI which are the same as the ones of the MPI table.
+As additional information regarding the SNPs another database was consulted, the UCSC database.
+This table contains every SNP of the MPI table, but with infromation about the reference genome and the locations of the SNPs.
+Having that information, it is possible to create statistics regarding the substitutions of the nucleotides, locations and the phenotypic results arising from those SNPs.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\linewidth]{pics/database.png}
+	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
+ 	\caption{\footnotesize{\textbf{Database}. 
+	~~~~~~~\\
+	The database created with information of MGI, MPI and UCSC.}}
+	\label{fig:database}
+\end{figure}
 

### 2015-03-09 article/article.tex
db Bild

--- 2015-03-09 article/article.tex
@@
 
-\Abstract{\textit{Nuclear receptors (NRs) are important signaling mediators in cells due to their ability to act as receptors and transcription factors.
-These molecules are activated by ligands and regulate critical functions in cell control, inflammation, fibrosis and tumor formation.
-Nuclear receptors act on the metabolism and signaling of cells by changing the expression of target genes.
-Here, we will investigate the knockout phenotypes and genetic variation of mouse NRs.
+\Abstract{\textit{
+Nuclear receptors (NRs) are a large family of ligand-activated transcription factors, that bind directly to DNA to regulate the expression of target genes.
+They regulate critical functions in cell control, inflammation, fibrosis and tumor formation and are involved in metabolism, development and reproduction.
+Nuclear receptors act on the metabolism and signalling of cells by changing the expression of target genes.
+Thus, they are associated with numerous pathologies such as cancer, cardiovascular disease, inflammation, and reproductive abnormalities.
+This paper will investigate the knockout phenotypes and genetic variation of mouse NRs.
 We will first create an assembly of all known mouse SNPs in the vicinity of mouse NR genes (+-1Mbp up-or downstream).
@@
 \section{Database}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\linewidth]{pics/db.png}
+	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
+ 	\caption{\footnotesize{\textbf{MGI phenotypes}. 
+	~~~~~~~\\
+	Mouse Genome Informatics phenotype distribution over the nuclear receptors.}}
+	\label{fig:mgi_pheotypes_distribution}
+\end{figure}
+~~~~~~~\\
 

### 2015-03-09 analysisMPI_top10_strains_distribution.pl

@@ -1,0 +1,95 @@
+#!/usr/bin/perl
+
+use strict;
+use warnings;
+use diagnostics;
+use Getopt::Long qw(GetOptions);
+use Pod::Usage;
+
+require MODULES::PlotsExtract;
+require MODULES::PlotsGenerate;
+
+my ($outputFile, $debug, $help, $man);
+
+GetOptions (
+'outputFile=s' => \$outputFile,
+'d|debug' => \$debug,
+'h|help' => \$help,
+'m|man' => \$man
+);
+
+pod2usage( -verbose => 1 ) if $help;
+pod2usage( -verbose => 2 ) if $man;
+
+usage() unless defined $outputFile;
+
+sub usage{
+    print "Incorrect parameters! \n";
+    print "Usage: perl analysisMPI_top10_strains_distribution.pl --outputFile <path_to_plot_file> [--d|debug]\n";
+    print "       $0 --help \n";
+    print "       $0 --man \n";
+    exit;
+}
+
+#extract plot data
+my ($strains, $percentage, $count, $total_strains) = MODULES::PlotsExtract::get_analysisMPI_top10_strains_distribution();
+
+#transform plot data
+my $x_axis                  = $strains;
+my $y_axis                  = $percentage;
+my $xlab                    = "Strains";
+my $ylab                    = "Strain frequency and z-score for nuclear receptors";
+my $legend_count_strains    = "Strain occurrences: ".$total_strains;
+my $legend_count_nr         = "Nuclear receptors in mouse: 49";
+my $label                   = $count;
+
+#generate plots
+MODULES::PlotsGenerate::plotVerticalString($x_axis, $y_axis, $outputFile, $xlab, $ylab, $legend_count_strains, $legend_count_nr, $label);
+
+
+__END__
+
+=head1 NAME
+
+ analysisMGI_phenotypes_distribution.pl - creates MGI-Phenotypes distribution plot for the phenotypes found in the MGI database.
+ Please include absolute path in the arguements.
+ 
+ =head1 SYNOPSYS
+ 
+ analysisMGI_phenotypes_distribution.pl [OPTIONS]
+ Options:
+ -debug debug message
+ -help brief help message
+ -man full documentation
+
+ =head1 OPTIONS
+ 
+ =over 8
+ 
+ =item B<-h|--help>
+ 
+ Prints a brief help message and exits
+ 
[…]

### 2015-03-09 article/article.tex
article

--- 2015-03-09 article/article.tex
@@
 
-\Abstract{\textit{Nuclear receptors (NRs) are important signaling mediators in cells due to their ability to act as receptors and transcription factors.
-These molecules are activated by ligands and regulate critical functions in cell control, inflammation, fibrosis and tumor formation.
-Nuclear receptors act on the metabolism and signaling of cells by changing the expression of target genes.
-Here, we will investigate the knockout phenotypes and genetic variation of mouse NRs.
-We will first create an assembly of all known mouse SNPs in the vicinity of mouse NR genes (+-1Mbp up-or downstream).
-In a second step, phenotype information for genetic knockouts and genetic variation data will be included.
-Knockout phenotypes are available from the MGI database, while the Mouse Phenome Database provides SNPs from various mouse strains, which can be correlated to extreme phenotypes, measured in these mouse strains.
-The goal of this analysis is to find NR SNPs in mice that influence changes in biological parameters such as body weight, body fat and other phenotypic traits.
+\Abstract{\textit{
+Nuclear receptors (NRs) are a large family of ligand-activated transcription factors, that bind directly to DNA to regulate the expression of target genes.
+They regulate critical functions in cell control, inflammation, fibrosis and tumor formation and are involved in metabolism, development and reproduction.
+Nuclear receptors act on the metabolism and signalling of cells by changing the expression of target genes and are associated with numerous pathologies such as cancer, cardiovascular disease, inflammation, and reproductive abnormalities.
+This paper presents the investigative results of knockout phenotypes and genetic variation for mouse NRs.
+Based on an assembly of all known mouse SNPs in the vicinity of mouse NR genes, the phenotype information for genetic knockouts and genetic variation data was compiled from public databases.
+Knockout phenotypes were extracted from the Mouse Genome Informatics (MGI) database, while the Mouse Phenome Database (MPD) provides SNPs from various mouse strains, which can be correlated with extreme phenotypes measured in these mouse strains.
+The goal of this analysis is to find NR-associated SNPs in mice that influence changes in biological parameters such as body weight, body fat and other phenotypic traits.
 Furthermore, we will couple these findings to phenotypes observed in mice with a targeted or spontaneous mutation of the nuclear receptor and thus provide additional indication for a putative functionality of the investigated SNPs.}}
@@
 
-\subsection{Nuclear receptors}
+\subsection{Mouse nuclear receptors}
 49 nuclear receptors
@@
 	\centering
+	\includegraphics[width=\linewidth]{pics/db.png}
+	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
+ 	\caption{\footnotesize{\textbf{MGI phenotypes}. 
+	~~~~~~~\\
+	Mouse Genome Informatics phenotype distribution over the nuclear receptors.}}
+	\label{fig:mgi_pheotypes_distribution}
+\end{figure}
+~~~~~~~\\
+
+\begin{figure}[H]
+	\centering
 	\includegraphics[width=\linewidth]{pics/database.png}
@@
 \end{figure}
+
+>>>>>>> origin/master
 

### 2015-03-09 article/article.tex
commit

--- 2015-03-09 article/article.tex
@@
 \subsection{Nuclear receptors}
-49 nuclear receptors
+The dataset on which this examination was based upon contains 49 nuclear receptors of mouse~\cite{proteomic} as well as additional information regarding these receptors.
+To achieve the goal of associating the receptors with extreme phenotypes, the dataset was analyzed and two mouse databases were brought in.
 \subsection{human genome}
@@
 Mouse Genome Informatics\footnote{\url{http://www.informatics.jax.org/}, \today} is a database for the laboratory mouse, which makes information about integrated genetics and associated phenotypes with their alleles available.
+
 For this report the MGI database was solely used for building a connection between the nuclear receptor genes and the associated phenotypes always dependent on miscellaneous strains.
@@
 Mouse Phenome Database\footnote{\url{http://phenome.jax.org/}, \today}~\cite{mpd} includes information about measured data on laboratoy mouse strains and populations as well as SNPs and phenotypes of the examined strains.
-The information about the SNPs and the variations were the motifs to also take a look at this database, wit
+More than 1330 were examined in several ways, over 3500 phenotype and 1.8 billion genotype annotations were made ~\cite{mpd}.
 Given that the MGI phenotypes can be compared to the ones found in MPI which are by far more detailed.
@@
 	\centering
-	\includegraphics[width=\linewidth]{pics/database.png}
+	\includegraphics[width=\linewidth]{pics/db.png}
 	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
@@
 	~~~~~~~\\
-	The database created with information of MGI, MPI and UCSC.}}
-	\label{fig:database}
+	Database scheme for the 49 nuclear receptors with information about MGI, MPI and UCSC.}}
+	\label{fig:mgi_pheotypes_distribution}
 \end{figure}
+~~~~~~~\\
+All databases were build upon the \textit{nr\_mapping} table, which contains the important association between the nuclear receptors and the gene names.
+Given that the table for MGI was created by searching the gene names in the database and extracting all necessary information, analogous for the MPI database.
 

### 2015-03-09 article/article.tex
abstract + intro

--- 2015-03-09 article/article.tex
@@
 They regulate critical functions in cell control, inflammation, fibrosis and tumor formation and are involved in metabolism, development and reproduction.
-Nuclear receptors act on the metabolism and signalling of cells by changing the expression of target genes and are associated with numerous pathologies such as cancer, cardiovascular disease, inflammation, and reproductive abnormalities.
+Nuclear receptors influence the metabolism and signalling processes in the cells by changing the expression of target genes and are associated with numerous pathologies such as cancer, cardiovascular disease, and reproductive abnormalities.
 This paper presents the investigative results of knockout phenotypes and genetic variation for mouse NRs.
@@
 The goal of this analysis is to find NR-associated SNPs in mice that influence changes in biological parameters such as body weight, body fat and other phenotypic traits.
-Furthermore, we will couple these findings to phenotypes observed in mice with a targeted or spontaneous mutation of the nuclear receptor and thus provide additional indication for a putative functionality of the investigated SNPs.}}
+Furthermore, these findings will be coupled to phenotypes observed in mice with a targeted or spontaneous mutation of the nuclear receptor and thus provide additional indication for a putative functionality of the investigated SNPs.}}
 
@@
 
-IntroIntroIntro
+Looking across the evolutionary patterns between mouse and human, numerous research experiments and gene regulation studies have shown striking similarities regarding certain processes and systems in the two organisms.
+The mouse presents up to 95\% genome similarity to humans and is thus often being used as a model organism when investigating anatomical, physiological or genetical markers in humans.
+Practically, mice are small, have an accelerated life cycle and represent a cost-effective alternative to genetic research and drug development for human diseases.
+Also, the majority of the genes responsible for complex diseases are shared between mice and humans, enhancing the chances of successfully identifying patterns in mice which would reveal human disease phenotypes~\cite{intro1}.
+~~~~~~~\\
+~~~~~~~\\  
+This paper makes use of the publicly available data in the Mouse Genome Informatics database and Mouse Phenome Database, respectively, in order to highlight changes in various biological parameters in mice under the influence of the NR-associated SNPs.
+Furthermore, these findings can be mapped to genotype - phenotype associations in humans, for studying the human biology and disease.
 
@@
 	~~~~~~~\\
-	Mouse Genome Informatics phenotype distribution over the nuclear receptors.}}
-	\label{fig:mgi_pheotypes_distribution}
-\end{figure}
-~~~~~~~\\
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\linewidth]{pics/database.png}
-	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
- 	\caption{\footnotesize{\textbf{Database}. 
-	~~~~~~~\\
 	The database created with information of MGI, MPI and UCSC.}}
@@
 \end{figure}
-
->>>>>>> origin/master
 

### 2015-03-10 article/article.tex
commit

@@
 
+\subsection{Nuclear receptors}
+The dataset on which this examination was based upon contains 49 nuclear receptors of mouse~\cite{proteomic} as well as additional information regarding these receptors.
+To achieve the goal of associating the receptors with extreme phenotypes, the dataset was analyzed and two mouse databases were brought in.
 \subsection{Mouse nuclear receptors}
@@
 Mouse Genome Informatics\footnote{\url{http://www.informatics.jax.org/}, \today} is a database for the laboratory mouse, which makes information about integrated genetics and associated phenotypes with their alleles available.
-For this report the MGI database was solely used for building a connection between the nuclear receptor genes and the associated phenotypes always dependent on miscellaneous strains.
-Information about SNPs and variation of the nuclear receptors were not taken from MGI, because the annotation regarding this matter was far better in the MPI and UCSC database.
+The Mouse Genome Database contains over 24000 genes with protein sequences and approximately 48000 genotypes with phenotype annotation ~\cite{mgd}.
+For this paper the MGD was solely used for building a connection between the nuclear receptor genes and the associated phenotypes always dependent on miscellaneous strains.
+Information about SNPs and variation of the nuclear receptors were not taken from MGD, because the annotation regarding this matter was far better in the MPI and UCSC database.
 \subsection{MPD}
 Mouse Phenome Database\footnote{\url{http://phenome.jax.org/}, \today}~\cite{mpd} includes information about measured data on laboratoy mouse strains and populations as well as SNPs and phenotypes of the examined strains.
-The information about the SNPs and the variations were the motifs to also take a look at this database, wit
+More than 1330 strains were examined in several ways, over 3500 phenotype and 1.8 billion genotype annotations were made~\cite{mpd}.
 Given that the MGI phenotypes can be compared to the ones found in MPI which are by far more detailed.
@@
 The goal of the database was to create several tables which could be mapped with the reference table of the 49 nuclear receptors, to create statistics about the relationship between the nuclear receptors and the phenotypes and SNPs associated with them.
-As mentioned above, the main problem was the connection between the MGI and the MPI tables, which was gone around by using an extra table which contained only the strains of MGI which are the same as the ones of the MPI table.
+As mentioned above, the main problem was the connection between the MGD and the MPD tables, which was gone around by using an extra table which contained only the strains of MGI which are the same as the ones of the MPD table.
 As additional information regarding the SNPs another database was consulted, the UCSC database.
-This table contains every SNP of the MPI table, but with infromation about the reference genome and the locations of the SNPs.
+This table contains every SNP of the MPD table, but with infromation about the reference genome and the locations of the SNPs.
 Having that information, it is possible to create statistics regarding the substitutions of the nucleotides, locations and the phenotypic results arising from those SNPs.
@@
 	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
- 	\caption{\footnotesize{\textbf{MGI phenotypes}. 
+ 	\caption{\footnotesize{\textbf{Database}. 
 	~~~~~~~\\
-	The database created with information of MGI, MPI and UCSC.}}
-	\label{fig:database}
+	Database scheme for the 49 nuclear receptors with information about MGI, MPI and UCSC.}}
+	\label{fig:mgi_pheotypes_distribution}
 \end{figure}
-
+~~~~~~~\\
+All databases were build upon the \textit{nr\_mapping} table, which contains the important association between the nuclear receptors and the gene names.
+The \textit{mgi} table with the information regarding the genes itself like allele attributes and types was connected with the \textit{mgi\_phenotypes} table which stored the phenotypic information in association with the internal MGD ID.
+To map the nuclear receptors and the data of MGD on MPD we used the \textit{mgi\_strains} table which contained every available strain in mgi and built with that and the gene names of \textit{mgi} the \textit{mpi\_strainmeans} table.
+This table stores information about experimental results from strains like means, maximum values, minimum value and the zscore which will be an important indicator in the results chapter.
+Further the table includes an \textit{measurementsnumber} which was the key identifier for an element in the \textit{mpi\_strainmeans} and the \textit{mpi\_measnum} table which stores information about phenotypes associated with the regarded genes.
+To get a more view on the genotypic aspect besides the phenotypic, the \textit{mpi\_rs} and the \textit{ucsc\_variants} tables were built.
+They store the information about SNPs in the genes which are affected by the nuclear receptors.
+Given that the table for MGD was created by searching the gene names in the database and extracting all necessary information, analogous for the MPD database.
 %------------------------------------------------

### 2015-03-10 article/article.tex
changes

--- 2015-03-09 article/article.tex
@@
 
-\subsection{Mouse nuclear receptors}
-49 nuclear receptors
-\subsection{human genome}
-tbd
+Both the process of consolidating a genotype - phenotype map, as well as the subsequent analysis of extreme phenotypes observed in mice rely on four biological elements: mouse strains, gene names / symbols, nuclear receptors, SNPs.
+The correlation between the position of a nuclear receptor and various SNPs on a gene can result in a specific phenotype.
+Due to increased specificity rate of the phenotype associated with a particular nuclear receptor, different NR and SNPs associations will result in different phenotypes for the same gene.
+Similarly, diverging mouse strains present non-identical phenotypes for the genetic parameters.
+~~~~~~~\\
+~~~~~~~\\
+The analysis presented in this paper was based upon a dataset consisting of the 49 nuclear receptors of mouse~\cite{proteomic} and their associated genes (for a full list of gene names, please see Annex~\ref{an:annex1}).
 
@@
 \section{Tools}
+
 \subsection{MGI}
+
 Mouse Genome Informatics\footnote{\url{http://www.informatics.jax.org/}, \today} is a database for the laboratory mouse, which makes information about integrated genetics and associated phenotypes with their alleles available.
+~~~~~~~\\
+~~~~~~~\\
 For this report the MGI database was solely used for building a connection between the nuclear receptor genes and the associated phenotypes always dependent on miscellaneous strains.
 Information about SNPs and variation of the nuclear receptors were not taken from MGI, because the annotation regarding this matter was far better in the MPI and UCSC database.
+
 \subsection{MPD}
+
 Mouse Phenome Database\footnote{\url{http://phenome.jax.org/}, \today}~\cite{mpd} includes information about measured data on laboratoy mouse strains and populations as well as SNPs and phenotypes of the examined strains.
-The information about the SNPs and the variations were the motifs to also take a look at this database, wit
+More than 1330 were examined in several ways, over 3500 phenotype and 1.8 billion genotype annotations were made ~\cite{mpd}.
 Given that the MGI phenotypes can be compared to the ones found in MPI which are by far more detailed.
@@
 \section{Database}
+
 The goal of the database was to create several tables which could be mapped with the reference table of the 49 nuclear receptors, to create statistics about the relationship between the nuclear receptors and the phenotypes and SNPs associated with them.
@@
 	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
- 	\caption{\footnotesize{\textbf{MGI phenotypes}. 
+ 	\caption{\footnotesize{\textbf{Database}. 
 	~~~~~~~\\
-	The database created with information of MGI, MPI and UCSC.}}
-	\label{fig:database}
+	Database scheme for the 49 nuclear receptors with information about MGI, MPI and UCSC.}}
+	\label{fig:mgi_pheotypes_distribution}
 \end{figure}
-
+~~~~~~~\\
+All databases were build upon the \textit{nr\_mapping} table, which contains the important association between the nuclear receptors and the gene names.
+Given that the table for MGI was created by searching the gene names in the database and extracting all necessary information, analogous for the MPI database.
 %------------------------------------------------
@@
 %----------------------------------------------------------------------------------------
+% Appendix
+%----------------------------------------------------------------------------------------
+\section{Annex1}
+\label{an:annex1}
+List of gene names associated with the 49 mouse nuclear receptors:
+~~~~~~~\\
+~~~~~~~\\
+\textit{Esrrg, Rorc, Nr1i2, Hnf4g, Nr0b1, Nr3c1, Hnf4a, Rxrg, Esrra, Thrb, Nr5a1, Pparg, Nr4a1, Nr5a2, Nr1h3, Nr4a3, Nr1i3, Nr2f2, Nr4a2, Nr2f6, Esrrb, Ppard, Nr2e1, Vdr, Nr6a1, Nr1h5, Rarb, Nr1h2, Esr1, Rora, Rarg, Nr1h4, Rara, Nr3c2, Pgr, Ar, Rxrb, Thra, Esr2, Rorb, Ppara, Nr1d1, Nr0b2, Nr2e3, Nr2f1, Nr1d2, Nr2c1, Rxra, Nr2c2}.
+
+%----------------------------------------------------------------------------------------
 %	REFERENCE LIST

### 2015-03-10 article/article.tex
commit

--- 2015-03-10 article/article.tex
@@
 This paper presents the investigative results of knockout phenotypes and genetic variation for mouse NRs.
-Based on an assembly of all known mouse SNPs in the vicinity of mouse NR genes, the phenotype information for genetic knockouts and genetic variation data was compiled from public databases.
+Based on an assembly of all known mouse SNPs in the mouse NR genes, the phenotype information for genetic knockouts and genetic variation data was compiled from public databases.
 Knockout phenotypes were extracted from the Mouse Genome Informatics (MGI) database, while the Mouse Phenome Database (MPD) provides SNPs from various mouse strains, which can be correlated with extreme phenotypes measured in these mouse strains.
@@
 
-\subsection{Nuclear receptors}
-The dataset on which this examination was based upon contains 49 nuclear receptors of mouse~\cite{proteomic} as well as additional information regarding these receptors.
-To achieve the goal of associating the receptors with extreme phenotypes, the dataset was analyzed and two mouse databases were brought in.
+Both the process of consolidating a genotype - phenotype map, as well as the subsequent analysis of extreme phenotypes observed in mice rely on four biological elements: mouse strains, gene names / symbols, nuclear receptors, SNPs.
+The correlation between the position of a nuclear receptor and various SNPs on a gene can result in a specific phenotype.
+Due to increased specificity rate of the phenotype associated with a particular nuclear receptor, different NR and SNPs associations will result in different phenotypes for the same gene.
+Similarly, diverging mouse strains present non-identical phenotypes for the genetic parameters.
+
 \subsection{Mouse nuclear receptors}
-49 nuclear receptors
-\subsection{human genome}
-tbd
+The analysis presented in this paper was based upon a dataset consisting of the 49 nuclear receptors of mouse~\cite{proteomic} and their associated genes (for a full list of gene names, please see Appendix~\ref{an:appendix}).
+The gene names were further on used to compile additional information regarding the position, associated SNPs and phenotypes for the nuclear receptors in mice.
 
-%------------------------------------------------
+\subsection{MGI - Mouse Genome Informatics}
 
-\section{Tools}
-\subsection{MGI}
+Mouse Genome Informatics\footnote{\url{http://www.informatics.jax.org/}, \today}~\cite{mgi} is free, online database for the laboratory mouse, which provides access to information about integrated genetics and associations between specific phenotypes and their corresponding alleles.
+~~~~~~~\\
+~~~~~~~\\
+For this research, the MGI database was solely used for building a connection between the nuclear receptor genes in the mouse and the associated phenotypes dependent on miscellaneous strains.
+Information about the SNPs and the variation of the nuclear receptors were not taken from MGI, because the annotation regarding this matter was far better in the MPI and UCSC database.
+
+\subsection{MPD - Mouse Phenome Database}
+
+Mouse Phenome Database\footnote{\url{http://phenome.jax.org/}, \today}~\cite{mpd} includes information about measured data on laboratoy mouse strains and populations as well as SNPs and phenotypes of the examined strains.
+More than 1330 were examined in several ways, over 3500 phenotype and 1.8 billion genotype annotations were made ~\cite{mpd}.
+Given that the MGI phenotypes can be compared to the ones found in MPI which are by far more detailed.
+The drawback of the MPI database is that every sort of data is always reffered on the regarding strain, making the mapping between the nuclear receptors and the MGI data more difficult.
 Mouse Genome Informatics\footnote{\url{http://www.informatics.jax.org/}, \today} is a database for the laboratory mouse, which makes information about integrated genetics and associated phenotypes with their alleles available.
@@
 \section{Database}
+
+The goal of the database was to create several tables which could be mapped with the reference table of the 49 nuclear receptors, to create statistics about the relationship between the nuclear receptors and the phenotypes and SNPs associated with them.
+As mentioned above, the main problem was the connection between the MGI and the MPI tables, which was gone around by using an extra table which contained only the strains of MGI which are the same as the ones of the MPI table.
+As additional information regarding the SNPs another database was consulted, the UCSC database.
+This table contains every SNP of the MPI table, but with infromation about the reference genome and the locations of the SNPs.
+Having that information, it is possible to create statistics regarding the substitutions of the nucleotides, locations and the phenotypic results arising from those SNPs.
 The goal of the database was to create several tables which could be mapped with the reference table of the 49 nuclear receptors, to create statistics about the relationship between the nuclear receptors and the phenotypes and SNPs associated with them.
@@
 ~~~~~~~\\
+All databases were build upon the \textit{nr\_mapping} table, which contains the important association between the nuclear receptors and the gene names.
+Given that the table for MGI was created by searching the gene names in the database and extracting all necessary information, analogous for the MPI database.
 All databases were build upon the \textit{nr\_mapping} table, which contains the important association between the nuclear receptors and the gene names.
@@
 
+\begin{table}[H]
+\centering
+\begin{tabulary}{\linewidth}{ LLLLLL }
+\hline
+	\textbf{Strain} & \textbf{Gene}  & \textbf{Feature} & \textbf{Phenotype} & \textbf{Z-score} & \textbf{Occurrence} \\  
+	\hline \hline 
+	CAST/EiJ  & Rora (RAR-related orphan receptor alpha) & protein coding gene & trigonelline relative abundance &  8.870 & 2440\\ 
+	C58/J  & Rorc (RAR-related orphan receptor gamma) & protein coding gene & N-acetylglutamate relative abundance & 8.450 & 1398\\ 
+	BUB/BnJ &  & 3-dehydrocarnitine  relative abundance &  8.420 & 1716\\ 
+	NOD/ShiLtJ   &  & glucose &  7.840 & 2599\\ 
+	NZW/LacJ   &  & glycerate  relative abundance &  7.410 & 2230\\ 
+	FVB/NJ  &  & succinylcarnitine  relative abundance  &  6.450 & 3335\\ 
+	NZB/BlNJ  &  & cornea-lens synechia  &  6.250 & 1342\\ 
+	SEA/GnJ  &  & optic cup changes &  6.170 & 912\\ 
+	NZO/HlLtJ &  &  body weight &  6.130 & 1007\\ 
+	129S1/SvImJ  &  & latency to reach periphery 10 min test  &  6.080 & 3360\\ 
+	\hline
+\end{tabulary}
+\caption{\footnotesize{\textbf{Top 10 extreme phenotypes.} 
+~~~~~~~\\
+Mouse strains associated with extreme phenotypes, based on their Z-score.}} 
+\label{tab:top10pheno}
+\end{table}
+~~~~~~~\\
+
 \begin{figure}[H]
@@
 	Mouse Phenome Database phenotype occurrence frequency among the strains associated with the nuclear receptors, having Z-score $>$ 2}}
-	\label{fig:mgi_pheotypes_strain_zscore2}
+	\label{fig:mpi_pheotypes_strain_zscore2}
 \end{figure}
@@
 	Mouse Phenome Database phenotype occurrence frequency among the strains associated with the nuclear receptors, having Z-score $<$ -2}}
-	\label{fig:mgi_pheotypes_strain_zscore2_neg}
+	\label{fig:mpi_pheotypes_strain_zscore2_neg}
 \end{figure}
@@
 %----------------------------------------------------------------------------------------
+% Appendix
+%----------------------------------------------------------------------------------------
+\section{Appendix}
+\label{an:appendix}
+List of gene names associated with the 49 mouse nuclear receptors:
+~~~~~~~\\
+~~~~~~~\\
+\textit{Esrrg, Rorc, Nr1i2, Hnf4g, Nr0b1, Nr3c1, Hnf4a, Rxrg, Esrra, Thrb, Nr5a1, Pparg, Nr4a1, Nr5a2, Nr1h3, Nr4a3, Nr1i3, Nr2f2, Nr4a2, Nr2f6, Esrrb, Ppard, Nr2e1, Vdr, Nr6a1, Nr1h5, Rarb, Nr1h2, Esr1, Rora, Rarg, Nr1h4, Rara, Nr3c2, Pgr, Ar, Rxrb, Thra, Esr2, Rorb, Ppara, Nr1d1, Nr0b2, Nr2e3, Nr2f1, Nr1d2, Nr2c1, Rxra, Nr2c2}.
+
+%----------------------------------------------------------------------------------------
 %	REFERENCE LIST

### 2015-03-10 article/article.tex
commit

--- 2015-03-10 article/article.tex
@@
 Similarly, diverging mouse strains present non-identical phenotypes for the genetic parameters.
+
+\subsection{Mouse nuclear receptors}
+
+The analysis presented in this paper was based upon a dataset consisting of the 49 nuclear receptors of mouse~\cite{proteomic} and their associated genes (242 unique gene Ids, 49 gene names / symbols); for a full list of gene names, see Appendix~\ref{an:appendix}).
+The gene names were further on used to compile additional information regarding the position, associated SNPs and phenotypes for the nuclear receptors in mice.
+
+\subsection{MGI - Mouse Genome Informatics}
+
+Mouse Genome Informatics\footnote{\url{http://www.informatics.jax.org/}, \today}~\cite{mgi} is a free, online database for the laboratory mouse and provides access to information about integrated genetics and associations between specific phenotypes and their corresponding alleles.
+It contains over 24000 genes and their protein sequences and approximately 48000 genotypes and phenotype annotations.
+For this research, the MGI database was solely used for building a connection between the nuclear receptor genes in the mouse and the associated phenotypes dependent on miscellaneous strains.
+
+\subsection{MPD - Mouse Phenome Database}
+
+Mouse Phenome Database\footnote{\url{http://phenome.jax.org/}, \today}~\cite{mpd} includes annotations of measured data on the laboratory mouse strains and populations, as well as SNPs and phenotypes of the examined strains.
+More than 1330 strains were examined, providing annotations for over 3500 phenotype and 1.8 billion genotypes.
 ~~~~~~~\\
 ~~~~~~~\\  
-The analysis presented in this paper was based upon a dataset consisting of the 49 nuclear receptors of mouse~\cite{proteomic} and their associated genes (for a full list of gene names, please see Annex~\ref{an:annex1}).
-
-%------------------------------------------------
-
-\section{Tools}
-
-\subsection{MGI}
-
-Mouse Genome Informatics\footnote{\url{http://www.informatics.jax.org/}, \today} is a database for the laboratory mouse, which makes information about integrated genetics and associated phenotypes with their alleles available.
-~~~~~~~\\
-~~~~~~~\\
-For this report the MGI database was solely used for building a connection between the nuclear receptor genes and the associated phenotypes always dependent on miscellaneous strains.
-Information about SNPs and variation of the nuclear receptors were not taken from MGI, because the annotation regarding this matter was far better in the MPI and UCSC database.
-
-\subsection{MPD}
-
-Mouse Phenome Database\footnote{\url{http://phenome.jax.org/}, \today}~\cite{mpd} includes information about measured data on laboratoy mouse strains and populations as well as SNPs and phenotypes of the examined strains.
-More than 1330 were examined in several ways, over 3500 phenotype and 1.8 billion genotype annotations were made ~\cite{mpd}.
-Given that the MGI phenotypes can be compared to the ones found in MPI which are by far more detailed.
-The drawback of the MPI database is that every sort of data is always reffered on the regarding strain, making the mapping between the nuclear receptors and the MGI data more difficult.
+The MPD database is more detailed and comprehensive than the MGI, so that the phenotypes found in MGI can be traced back to MPD.
+However, the MPD strictly associates individual phenotypes to their corresponding strain, posing difficulties in the mapping process between the mouse nuclear receptors and the MGI data.
 
@@
 
-The goal of the database was to create several tables which could be mapped with the reference table of the 49 nuclear receptors, to create statistics about the relationship between the nuclear receptors and the phenotypes and SNPs associated with them.
-As mentioned above, the main problem was the connection between the MGI and the MPI tables, which was gone around by using an extra table which contained only the strains of MGI which are the same as the ones of the MPI table.
-As additional information regarding the SNPs another database was consulted, the UCSC database.
-This table contains every SNP of the MPI table, but with infromation about the reference genome and the locations of the SNPs.
-Having that information, it is possible to create statistics regarding the substitutions of the nucleotides, locations and the phenotypic results arising from those SNPs.
-
+Since annotation from multiple public databases and internal data were used, there was a need to design and implement a database supporting the research.
+The database structure thereby presents 7 tables and aims to facilitate the processes of gathering, parsing and statistical analysis of the nuclear receptors data, as well as to allow for a clear mapping between the nuclear receptor genes and the phenotypes and SNPs associated with them.
+The main challenge was to establish a connection between the MGI, MPD and UCSC tables.
+Therefore, there are 2 tables containing information from the MGI, 3 tables for MPD, 1 USCS table and the core table:
+\begin{itemize}
+\item \textbf{nr\_mapping} - core table, which contains the associations between the 49 nuclear receptors in the mouse and their corresponding gene names. 
+\item \textbf{mgi} - contains the MGI gene annotations (MGI gene id, type, attributes, transmission etc.). 
+\item \textbf{mgi\_phenotypes} - stores phenotypic information in association with MGI genes; more than one phenotype can be associated with a specific gene.
+\item \textbf{mpd\_snp} - contains the MPD SNP-annotations in correlation with the nuclear receptor genes. 
+\item \textbf{ucsc} - contains a more detailed overview on the MPD mutations from \textit{mpd\_snp} 
+\item \textbf{mpd\_strains} - contains the MPD strain annotations (MPD strain id, sex, number of mice, Z-score etc.)
+\item \textbf{mpd\_phenotypes} - stores phenotypic information in association with MPD strains; more than one phenotype can be associated with a specific strain.
+\end{itemize} 
+~~~~~~~\\  
 \begin{figure}[H]
@@
 \end{figure}
-~~~~~~~\\
-All databases were build upon the \textit{nr\_mapping} table, which contains the important association between the nuclear receptors and the gene names.
-Given that the table for MGI was created by searching the gene names in the database and extracting all necessary information, analogous for the MPI database.
 %------------------------------------------------
@@
 \section{Results}
+The goal hereby was to visualise the most important phenotypes associated with the 49 nuclear receptors in the mouse.
+In this regard, the phenotypic and genomic data from the MGI and MPD was compared and analysed.
+All graphics were generated using the R package\footnote{\url{www.r-project.org/}, \today}.
+
+\subsection{MGI Statistics}
 
@@
 	~~~~~~~\\
-	Mouse Genome Informatics phenotype distribution over the nuclear receptors.}}
+	Mouse Genome Informatics phenotype distribution over the genes associated with the 49 nuclear receptors in the mouse.}}
 	\label{fig:mgi_pheotypes_distribution}
@@
 ~~~~~~~\\
+The MGI database consists of generalised definitions of the phenotypes associated with the nuclear receptors found in the mouse.
+Figure~\ref{fig:mgi_pheotypes_distribution} illustrates the most significant phenotypes and their occurrence frequency across the 242 genes corresponding to the nuclear receptors in the mouse.
+With a count of 196 matches across the dataset, the most prominent phenotype correlates with the homeostatic metabolic processes, such as temperature regulation and pH-balance.
+Moreover, Figure~\ref{fig:mgi_pheotypes_nr} provides an insight into the exact association of each phenotype with the corresponding nuclear receptor genes, such that homeostatis, for instance, is prominently found in the following genes: \textit{Esr1}, \textit{Pparg}, \textit{Thrb}, \textit{Thra}, \textit{Vdr} etc.
+Other phenotypes describe body size and growth features and are representative for the \textit{Pparg} genes.
 
-%Figure~\ref{fig:mgi_pheotypes_distribution} bla bla 
 \begin{figure}[H]
 	\centering
-	\includegraphics[width=\linewidth]{pics/mgi_types_distribution.pdf}
+	\includegraphics[width=\linewidth, height=0.99\linewidth]{pics/mgi_phenotypes_nr.pdf}
 	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
- 	\caption{\footnotesize{\textbf{MGI allele types}. 
+ 	\caption{\footnotesize{\textbf{MGI phenotype - nuclear receptor gene associations}. 
 	~~~~~~~\\
-	Mouse Genome Informatics allele type distribution over the nuclear receptors.}}
-	\label{fig:mgi_types_distribution}
+	Mouse Genome Informatics phenotype occurrence frequency among the nuclear receptor genes.}}
+	\label{fig:mgi_pheotypes_nr}
 \end{figure}
@@
 
+\subsection{MPD Statistics}
 \begin{figure}[H]
@@
 	~~~~~~~\\
-	Mouse Phenotype Database phenotype distribution over the nuclear receptors, having Z-score $>$ 2}}
+	Mouse Phenotype Database extreme phenotype distribution over the genes associated with the 49 nuclear receptors in the mouse, having Z-score $>$ 2}}
 	\label{fig:mpi_pheotypes_distribution_zscore2}
@@
 	~~~~~~~\\
-	Mouse Phenotype Database phenotype distribution over the nuclear receptors, having Z-score $<$ -2}}
+	Mouse Phenotype Database extreme phenotype distribution over the genes associated with the 49 nuclear receptors in the mouse, having Z-score $<$ -2}}
 	\label{fig:mpi_pheotypes_distribution_zscore2_neg}
+\end{figure}
+~~~~~~~\\
+~~~~~~~\\
+The MPDatabase contains annotations for extreme phenotypes associated with different mouse strains.
+An extreme phenotype is described by a Z-score above 2 (very high) or below -2 (very low), relative to other measurement means.
+Based on these specifications, Figure~\ref{fig:mpi_pheotypes_distribution_zscore2} illustrates the most significant phenotypes and their occurrence frequency across the 242 genes corresponding to the nuclear receptors in the mouse, with a Z-score $>$ 2 (extreme high phenotypes).
+Here, the most significant phenotypes include \textit{body weight}, \textit{kidney weight} and \textit{cholesterol}.
+Similarly, Figure~\ref{fig:mpi_pheotypes_distribution_zscore2_neg} presents the most significant phenotypes and their occurrence frequency across the 242 genes corresponding to the nuclear receptors in the mouse, having a Z-score $<$ -2 (extreme low phenotypes).
+In this case, the most significant phenotypes include \textit{body weight} as well, but the focus lies on several blood phenotypes (e.g.
+lymphocyte differential, red blood cell count, hemoglobin).
+~~~~~~~\\
+~~~~~~~\\
+Table~\ref{tab:top10pheno} shows the strains and gene names associated with the most significant MPD phenotypes. 
+\begin{table}[H]
+\centering
+\begin{tabulary}{\linewidth}{ LLLLLL }
+\hline
+	\textbf{Strain} & \textbf{Gene}  & \textbf{Feature} & \textbf{Phenotype} & \textbf{Z-score} & \textbf{Occurrence} \\  
+	\hline \hline 
+	CAST/EiJ  & Rora (RAR-related orphan receptor alpha) & protein coding gene & trigonelline relative abundance &  8.870 & 2440\\ 
+	C58/J  & Rorc (RAR-related orphan receptor gamma) & protein coding gene & N-acetylglutamate relative abundance & 8.450 & 1398\\ 
+	BUB/BnJ &  & 3-dehydrocarnitine  relative abundance &  8.420 & 1716\\ 
+	NOD/ShiLtJ   &  & & glucose &  7.840 & 2599\\ 
+	NZW/LacJ   &  & & glycerate relative abundance &  7.410 & 2230\\ 
+	FVB/NJ  &  & & succinylcarnitine relative abundance  &  6.450 & 3335\\ 
+	NZB/BlNJ  &  & & cornea-lens synechia  &  6.250 & 1342\\ 
+	SEA/GnJ  &  & & optic cup changes &  6.170 & 912\\ 
+	NZO/HlLtJ & & & body weight &  6.130 & 1007\\ 
+	129S1/SvImJ  & & & latency to reach periphery 10 min test  &  6.080 & 3360\\ 
+	\hline
+\end{tabulary}
+\caption{\footnotesize{\textbf{Top 10 extreme phenotypes.} 
+~~~~~~~\\
+Mouse strains associated with extreme phenotypes, based on their Z-score.}} 
+\label{tab:top10pheno}
+\end{table}
+~~~~~~~\\
+Moreover, Figures~\ref{fig:mpi_pheotypes_strain_zscore2} and~\ref{fig:mpi_pheotypes_strain_zscore2_neg} provide an insight into the exact association of each phenotype with the corresponding nuclear receptor genes.
+Therefore, there are ...
+genes associated with the extreme high phenotypes and 5 genes associated with the extreme low phenotypes, respectively, as following: \textit{Rora}, \textit{Esr1}, \textit{Esrrg}, \textit{Thrb} and \textit{Nr3c2}.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\linewidth]{pics/mpi_phenotypes_nr_zscore2.pdf}
+	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
+ 	\caption{\footnotesize{\textbf{MPD extreme phenotypes, having Z-score $>$ 2}. 
+	~~~~~~~\\
+	Mouse Phenome Database extreme phenotype occurrence frequency among the 49 nuclear receptors in the mouse, having Z-score $>$ 2}}
+	\label{fig:mpi_pheotypes_strain_zscore2}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=\linewidth]{pics/mgi_phenotypes_nr.pdf}
+	\includegraphics[width=\linewidth]{pics/mpi_phenotypes_nr_zscore2_neg.pdf}
 	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
- 	\caption{\footnotesize{\textbf{MGI phenotype - nuclear receptor associations}. 
+ 	\caption{\footnotesize{\textbf{MPD extreme phenotypes, having Z-score $<$ -2}. 
 	~~~~~~~\\
-	Mouse Genome Informatics phenotype occurrence frequency among the nuclear receptors.}}
-	\label{fig:mgi_pheotypes_nr}
+	Mouse Phenome Database extreme phenotype occurrence frequency among the 49 nuclear receptors in the mouse, having Z-score $<$ -2}}
+	\label{fig:mpi_pheotypes_strain_zscore2_neg}
 \end{figure}
 ~~~~~~~\\
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\linewidth]{pics/mgi_types_nr.pdf}
-	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
- 	\caption{\footnotesize{\textbf{MGI allele type - nuclear receptor associations}. 
-	~~~~~~~\\
-	Mouse Genome Informatics allele type occurrence frequency among the nuclear receptors.}}
-	\label{fig:mgi_types_nr}
-\end{figure}
-~~~~~~~\\
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\linewidth]{pics/mpi_phenotypes_strain_zscore2.pdf}
-	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
- 	\caption{\footnotesize{\textbf{MPD phenotype - strain associations, having Z-score $>$ 2}. 
-	~~~~~~~\\
-	Mouse Phenome Database phenotype occurrence frequency among the strains associated with the nuclear receptors, having Z-score $>$ 2}}
-	\label{fig:mgi_pheotypes_strain_zscore2}
-\end{figure}
-~~~~~~~\\
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\linewidth]{pics/mpi_phenotypes_strain_zscore2.pdf}
-	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
- 	\caption{\footnotesize{\textbf{MPD phenotype - strain associations, having Z-score $<$ -2}. 
-	~~~~~~~\\
-	Mouse Phenome Database phenotype occurrence frequency among the strains associated with the nuclear receptors, having Z-score $<$ -2}}
-	\label{fig:mgi_pheotypes_strain_zscore2_neg}
-\end{figure}
 
@@
 %----------------------------------------------------------------------------------------
-\section{Annex1}
-\label{an:annex1}
+\section{Appendix}
+\label{an:appendix}
 List of gene names associated with the 49 mouse nuclear receptors:

### 2015-03-11 article/article.tex
commit

@@
 \hline
-	\textbf{Strain} & \textbf{Gene}  & \textbf{Feature} & \textbf{Phenotype} & \textbf{Z-score} & \textbf{Occurrence} \\  
+	\textbf{Strain} & \textbf{Gene}  & \textbf{Feature} & \textbf{Phenotype} & \textbf{Z-score}
+	\\  
 	\hline \hline 
-	CAST/EiJ  & Rora (RAR-related orphan receptor alpha) & protein coding gene & trigonelline relative abundance &  8.870 & 2440\\ 
-	C58/J  & Rorc (RAR-related orphan receptor gamma) & protein coding gene & N-acetylglutamate relative abundance & 8.450 & 1398\\ 
-	BUB/BnJ &  & 3-dehydrocarnitine  relative abundance &  8.420 & 1716\\ 
-	NOD/ShiLtJ   &  & & glucose &  7.840 & 2599\\ 
-	NZW/LacJ   &  & & glycerate relative abundance &  7.410 & 2230\\ 
-	FVB/NJ  &  & & succinylcarnitine relative abundance  &  6.450 & 3335\\ 
-	NZB/BlNJ  &  & & cornea-lens synechia  &  6.250 & 1342\\ 
-	SEA/GnJ  &  & & optic cup changes &  6.170 & 912\\ 
-	NZO/HlLtJ & & & body weight &  6.130 & 1007\\ 
-	129S1/SvImJ  & & & latency to reach periphery 10 min test  &  6.080 & 3360\\ 
+	CAST/EiJ  & Rora (RAR-related orphan receptor alpha) & protein coding gene & trigonelline relative abundance &  8.8 
+	C58/J  & Rorc (RAR-related orphan receptor gamma) & protein coding gene & N-acetylglutamate relative abundance & 8.4\\ 
+	CAST/EiJ  & Rara (retinoic acid receptor, alpha) & protein coding gene & N2-acetyllysine (16h fast), relative abundance &  4.6\\ 
+	NZB/BlNJ & Nr1i3 (nuclear receptor subfamily 1, group I, member 3) & protein coding gene & [ECG parameters]interval between peak of P-wave to R-wave (PR) &  4.6\\ 
+	CE/J & Nr0b1 (nuclear receptor subfamily 0, group B, member 1) & protein coding gene & kidney total weight &  4.4\\
+	NZB/BlNJ  & Nr5a1 (nuclear receptor subfamily 5, group A, member 1) & protein coding gene & succinylcarnitine relative abundance  &  3.8\\
+	C57L/J & Esr2 (estrogen receptor 2 beta) & protein coding gene & 
+	relative size of perivascular immune cell clusters &  3.5\\
+	NOD/ShiLtJ  & Ppard (peroxisome proliferator activator receptor delta)  & protein coding gene & 
+	percentage of parasites in brain relative to all organs tested &  3.1\\ 
+	BUB/BnJ & Esr1 (estrogen receptor 1 alpha) & protein coding gene & 3-dehydrocarnitine  relative abundance &  2.7\\  
 	\hline

### 2015-03-11 article/article.tex
presentation initial commit

@@
 \centering
-\begin{tabulary}{\linewidth}{ LLLLLL }
+\begin{tabulary}{\linewidth}{ LLLLL }
 \hline
@@
 	\hline \hline 
-	CAST/EiJ  & Rora (RAR-related orphan receptor alpha) & protein coding gene & trigonelline relative abundance &  8.8 
+	CAST/EiJ  & Rora (RAR-related orphan receptor alpha) & protein coding gene & trigonelline relative abundance &  8.8 \\
 	C58/J  & Rorc (RAR-related orphan receptor gamma) & protein coding gene & N-acetylglutamate relative abundance & 8.4\\ 
 	CAST/EiJ  & Rara (retinoic acid receptor, alpha) & protein coding gene & N2-acetyllysine (16h fast), relative abundance &  4.6\\ 
-	NZB/BlNJ & Nr1i3 (nuclear receptor subfamily 1, group I, member 3) & protein coding gene & [ECG parameters]interval between peak of P-wave to R-wave (PR) &  4.6\\ 
+	NZB/BlNJ & Nr1i3 (nuclear receptor subfamily 1, group I, member 3) & protein coding gene & ECG parameters interval between peak of P-wave to R-wave (PR) &  4.6\\ 
 	CE/J & Nr0b1 (nuclear receptor subfamily 0, group B, member 1) & protein coding gene & kidney total weight &  4.4\\

### 2015-03-11 article/article.tex
commit

@@
 \hline
-	\textbf{Strain} & \textbf{Gene}  & \textbf{Feature} & \textbf{Phenotype} & \textbf{Z-score}
+	\textbf{Strain} & \textbf{Gene}  & \textbf{Phenotype} & \textbf{Z-score}
 	\\  
 	\hline \hline 
-	CAST/EiJ  & Rora (RAR-related orphan receptor alpha) & protein coding gene & trigonelline relative abundance &  8.8 \\
-	C58/J  & Rorc (RAR-related orphan receptor gamma) & protein coding gene & N-acetylglutamate relative abundance & 8.4\\ 
-	CAST/EiJ  & Rara (retinoic acid receptor, alpha) & protein coding gene & N2-acetyllysine (16h fast), relative abundance &  4.6\\ 
-	NZB/BlNJ & Nr1i3 (nuclear receptor subfamily 1, group I, member 3) & protein coding gene & ECG parameters interval between peak of P-wave to R-wave (PR) &  4.6\\ 
-	CE/J & Nr0b1 (nuclear receptor subfamily 0, group B, member 1) & protein coding gene & kidney total weight &  4.4\\
-	NZB/BlNJ  & Nr5a1 (nuclear receptor subfamily 5, group A, member 1) & protein coding gene & succinylcarnitine relative abundance  &  3.8\\
-	C57L/J & Esr2 (estrogen receptor 2 beta) & protein coding gene & 
+	CAST/EiJ  & Rora (RAR-related orphan receptor alpha) & trigonelline relative abundance &  8.8 \\
+	C58/J  & Rorc (RAR-related orphan receptor gamma) & N-acetylglutamate relative abundance & 8.4\\ 
+	CAST/EiJ  & Rara (retinoic acid receptor, alpha) & N2-acetyllysine (16h fast), relative abundance &  4.6\\ 
+	NZB/BlNJ & Nr1i3 (nuclear receptor subfamily 1, group I, member 3) & ECG parameters interval between peak of P-wave to R-wave (PR) &  4.6\\ 
+	CE/J & Nr0b1 (nuclear receptor subfamily 0, group B, member 1) & kidney total weight &  4.4\\
+	NZB/BlNJ  & Nr5a1 (nuclear receptor subfamily 5, group A, member 1) & succinylcarnitine relative abundance  &  3.8\\
+	C57L/J & Esr2 (estrogen receptor 2 beta) & 
 	relative size of perivascular immune cell clusters &  3.5\\
-	NOD/ShiLtJ  & Ppard (peroxisome proliferator activator receptor delta)  & protein coding gene & 
+	NOD/ShiLtJ  & Ppard (peroxisome proliferator activator receptor delta) & 
 	percentage of parasites in brain relative to all organs tested &  3.1\\ 
-	BUB/BnJ & Esr1 (estrogen receptor 1 alpha) & protein coding gene & 3-dehydrocarnitine  relative abundance &  2.7\\  
+	BUB/BnJ & Esr1 (estrogen receptor 1 alpha) & 3-dehydrocarnitine  relative abundance &  2.7\\  
 	\hline

### 2015-03-11 article/article.tex
commit
presentation changes

@@
 ~~~~~~~\\
-Table~\ref{tab:top10pheno} shows the strains and gene names associated with the most significant MPD phenotypes. 
+Table~\ref{tab:top10pheno} shows the strains and gene names associated with the highest ranking MPD phenotypes relative to their Z-scores. 
 \begin{table}[H]
 \centering
-\begin{tabulary}{\linewidth}{ LLLLL }
+\begin{tabulary}{\linewidth}{ LLLL }
 \hline
@@
 	relative size of perivascular immune cell clusters &  3.5\\
-	NOD/ShiLtJ  & Ppard (peroxisome proliferator activator receptor delta) & 
+	NOD/\\ShiLtJ & Ppard (peroxisome proliferator activator receptor delta) & 
 	percentage of parasites in brain relative to all organs tested &  3.1\\ 

### 2015-03-12 article/article.tex
commit

--- 2015-03-11 article/article.tex
@@
 	Database scheme for the 49 nuclear receptors with information about MGI, MPI and UCSC.}}
-	\label{fig:mgi_pheotypes_distribution}
+	\label{fig:database}
 \end{figure}
@@
 ~~~~~~~\\
+
+%------------------------------------------------
+\section{The 49 nuclear receptors in the human genome}
+
+In contrast to the nuclear receptors in mouse, there have been two studies about nuclear receptors in human, called \textit{kora} and \textit{twinsuk}.
+With the results of these studies the database used in this paper was expanded beyond one species with the goal to find similarities between the nuclear receptors in human and mouse.
+Because of the lack of researched nuclear receptors in mouse there can be only made comparisons regarding the phenotype with the tables \textit{mgi\_phenotypes} \ref{fig:database} and \textit{mpg\_phenotypes} \ref{fig:database}.
+Table \ref{tab:compHuMo} shows a list with nuclear receptors which are found in mouse and human and have a similar phenotype or a similar biological function.
+~~~~~~~\\
+\begin{table}[H]
+\centering
+\begin{tabulary}{\linewidth}{ LLLLL }
+\hline
+	\textbf{Gene} & \textbf{Mouse phenotype}  & \textbf{Human phenotype}
+	\\  
+	\hline \hline 
+	Nr2e1  & corticosterone, \textit{steroid hormone} & lathosterol, \textit{steroid hormone}\\
+	Essrg  & carnitine, \textit{lysine metabolism} & glutaroyl carnitine, \textit{lysine metabolism}\\
+	Thrb  & ,\textit{thyroid hormone resistance} & \textit{thyroid hormone resistance}\\
+	Nr0b1 & \textit{adrenal hypoplasia} & \textit{adrenal hypoplasia}\\
+	Vdr & \textit{Osteoporosis} & \textit{Osteoporosis}\\
+	Ar & Spinal and bulbar muscular atrophy & Spinal and bulbar muscular atrophy\\
+	\hline
+\end{tabulary}
+\caption{\footnotesize{\textbf{Nuclear receptors with their coding protein or phenotype} 
+~~~~~~~\\
+Nuclear receptors in human and mouse which share a similar protein or a similar biological function.}} 
+\label{tab:compHuMo}
+\end{table}
 

### 2015-03-12 article/article.tex

@@
 ~~~~~~~\\
-Table~\ref{tab:top10pheno} shows the strains and gene names associated with the most significant MPD phenotypes. 
+Table~\ref{tab:top10pheno} shows the strains and gene names associated with the highest ranking \textit{MPD} phenotypes relative to their Z-scores.
+As seen in the table the phenotypes are not directly associated with expressed proteins rather than observed abnormal condition of mouse strains.
+Furthermore the denoted phenotypes in the table of \textit{MPD} show nevertheless a more detailed description than \textit{MGI}.
 \begin{table}[H]
 \centering
-\begin{tabulary}{\linewidth}{ LLLLL }
+\begin{tabulary}{\linewidth}{ LLLL }
 \hline
@@
 	relative size of perivascular immune cell clusters &  3.5\\
+	CE/J & Rorb (RAR-related orphan receptor beta) & 
+	tigloylglycine (16h fast), relative abundance&  3.3\\
 	NOD/\\ShiLtJ & Ppard (peroxisome proliferator activator receptor delta) & 
@@
 \end{figure}
-~~~~~~~\\
 
 %------------------------------------------------
-\section{The 49 nuclear receptors in the human genome}
+\section{Nuclear receptors in the human genome}
 
@@
 \begin{table}[H]
-\centering
 \begin{tabulary}{\linewidth}{ LLLLL }
@@
 	Essrg  & carnitine, \textit{lysine metabolism} & glutaroyl carnitine, \textit{lysine metabolism}\\
-	Thrb  & ,\textit{thyroid hormone resistance} & \textit{thyroid hormone resistance}\\
+	Thrb  & \textit{thyroid hormone resistance} & \textit{thyroid hormone resistance}\\
 	Nr0b1 & \textit{adrenal hypoplasia} & \textit{adrenal hypoplasia}\\
 	Vdr & \textit{Osteoporosis} & \textit{Osteoporosis}\\
-	Ar & Spinal and bulbar muscular atrophy & Spinal and bulbar muscular atrophy\\
+	Ar & \textit{Spinal and bulbar muscular atrophy} & \textit{Spinal and bulbar muscular atrophy}\\
 	\hline
@@
 
-Blabla
+The statistics about extreme phenotypes which manifest if a variation in a nuclear receptor in mice occurs, show clearly that very important fields are affected which leads to the conclusion that nuclear receptors have a big influence on the development of the mouse.
+The data of \textit{MGI} and \textit{MPD} on nuclear receptors is limited, the phenotypes annotated in \textit{MGI} decribe more the biological systems in which the phenotypes occur rather than the description of differences to a non variant like in \textit{MPD}.
+Nevertheless, in contrast to the human genome where there are less annotations than in the mouse databases, the conclusion on the comparison between the two species shows a shallow match regarding the phenotypic expression in case of a variation in one of the same nuclear receptors.
 

### 2015-03-12 article/article.tex
discussion

--- 2015-03-11 article/article.tex
@@
 	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
- 	\caption{\footnotesize{\textbf{MPD phenotypes, having Z-score $>$ 2}. 
+ 	\caption{\footnotesize{\textbf{MPD extreme high phenotypes, having Z-score $>$ 2}. 
 	~~~~~~~\\
-	Mouse Phenotype Database extreme phenotype distribution over the genes associated with the 49 nuclear receptors in the mouse, having Z-score $>$ 2}}
+	Mouse Phenotype Database extreme high phenotype distribution over the genes associated with the 49 nuclear receptors in the mouse, having Z-score $>$ 2}}
 	\label{fig:mpi_pheotypes_distribution_zscore2}
@@
 	~~~~~~~\\
-	Mouse Phenotype Database extreme phenotype distribution over the genes associated with the 49 nuclear receptors in the mouse, having Z-score $<$ -2}}
+	Mouse Phenotype Database extreme low phenotype distribution over the genes associated with the 49 nuclear receptors in the mouse, having Z-score $<$ -2}}
 	\label{fig:mpi_pheotypes_distribution_zscore2_neg}
@@
 ~~~~~~~\\
-Table~\ref{tab:top10pheno} shows the strains and gene names associated with the highest ranking MPD phenotypes relative to their Z-scores. 
+Table~\ref{tab:top10pheno} shows the strains and gene names associated with the highest ranking \textit{MPD} phenotypes relative to their Z-scores.
+As seen in the table, the phenotypes are not directly associated with the expressed proteins, but rather with the observed abnormal condition in the mouse strains.
+%Furthermore the denoted phenotypes in the table of \textit{MPD} show nevertheless a more detailed description than \textit{MGI}.
 \begin{table}[H]
@@
 ~~~~~~~\\
-Mouse strains associated with extreme phenotypes, based on their Z-score.}} 
+Mouse strains associated with MPD extreme phenotypes, based on their Z-score.}} 
 \label{tab:top10pheno}
@@
 \end{figure}
+
+%------------------------------------------------
+\subsection{Nuclear receptors in the human genome}
+The nuclear receptor variation has been widely studied not only in the mouse, but also in the human genome.
+The research hereby refers to two published genome-wide association studies in this field of interest, namely \textit{Kora}\footnote{\url{http://epi.helmholtz-muenchen.de/kora-gen}, \today}  - a GWAS study initiated at the Helmholtz Zentrum in Munich involving approximately 18,000 adults from Southern Germany - and \textit{TwinsUK}\footnote{\url{http://www.twinsuk.ac.uk/}, \today}, being the biggest adult twin genetic registry in the UK.
 ~~~~~~~\\
+~~~~~~~\\
+Using the information provided by these studies regarding the correlations between different SNPs and metabolites to metabolic pathways, this research aimed to identify the similarities between the nuclear receptors and their associated phenotypes in human and in the mouse, respectively.
+Due to missing or incomplete annotations for the nuclear receptors in the mouse, comparisons could only be made between the phenotypes extracted from the MGI and MPD databases (see tables \textit{mgi\_phenotypes} and \textit{mpg\_phenotypes} from the internal database~\ref{fig:database}).
+Table \ref{tab:compHuMo} shows a list of nuclear receptors found in both the mouse and human and which present a similar phenotype or share the same biological function.
+~~~~~~~\\
+~~~~~~~\\
+%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%
+%which are the joined phenotypes, what's the consequence? e.g. what does it mean that mouse + human share steroid hormone? 
+%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%
+~~~~~~~\\
+\begin{table}[H]
+\begin{tabulary}{\linewidth}{ LLL }
+\hline
+	\textbf{Nuclear Receptor Gene} & \textbf{Mouse phenotype}  & \textbf{Human phenotype}
+	\\  
+	\hline \hline 
+	Nr2e1  & corticosterone, \textit{steroid hormone} & lathosterol, \textit{steroid hormone}\\
+	Essrg  & carnitine, \textit{lysine metabolism} & glutaroyl carnitine, \textit{lysine metabolism}\\
+	Thrb  & \textit{thyroid hormone resistance} & \textit{thyroid hormone resistance}\\
+	Nr0b1 & \textit{adrenal hypoplasia} & \textit{adrenal hypoplasia}\\
+	Vdr & \textit{osteoporosis} & \textit{osteoporosis}\\
+	Ar & \textit{spinal and bulbar muscular atrophy} & \textit{spinal and bulbar muscular atrophy}\\
+	\hline
+\end{tabulary}
+\caption{\footnotesize{\textbf{Nuclear receptors with their coding protein or phenotype} 
+~~~~~~~\\
+Nuclear receptors in human and in the mouse which share a similar phenotype or biological function.}} 
+\label{tab:compHuMo}
+\end{table}
 
@@
 
-Blabla
+Over the last year, the number of studies and publications involving the nuclear receptor variation has increased and scientist have been focusing on exploring the effects of nuclear receptors in different organisms, on various levels: in the cells, tissues, and even full-body impact.
+The goal hereby is to link their metabolic activity to possible human diseases and facilitate the drug development processes.
+In this regard, extensive experiments and studies have been performed on humans, as well as on laboratory mice, which are known to present a striking genomic resemblance to humans (up to 95\%) and and time and money-wise more advantageous.
+As a result, it has been proved that nuclear receptors have a great influence over the extreme phenotypes showcased in the mouse, as well as in human.
+~~~~~~~\\
+~~~~~~~\\
+In spite of the ascending progress in this field of research, there annotation data provided in the mouse public databases - \textit{Mouse Genome Informatics} and \textit{Mouse Phenome Database} on nuclear receptors is still rather limited.
+However, the two databases provide complementary information: the phenotypes annotated in \textit{MGI} describe roughly the biological systems in which they occur rather than the description of differences between non-variants, as of \textit{MPD}.
+On the other hand, the UCSC database for the human genome is far more detailed, comprising also information regarding the metabolic pathways associated with various phenotypes.
+~~~~~~~\\
+~~~~~~~\\
+This research presents the most significant phenotypes found in the mouse, based on the 49 nuclear receptor variation, as well as a mapping of these phenotypes on the human phenotypic expression, providing insight into possible human disease conditions.
+Nevertheless, the mouse public databases are far less comprehensive than the UCSC, for instance, allowing for only a generalised comparison between the two species, relative to the phenotypic expression in case of a variation in one of the same nuclear receptors.
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%what are the similarites, differences? How does the mouse study involve human phenotypes?
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2015-03-13 article/article.tex
commit

@@
 Table \ref{tab:compHuMo} shows a list of nuclear receptors found in both the mouse and human and which present a similar phenotype or share the same biological function.
+As seen there are only a few phenotypes shared between these two species.
+The reason therefore is that the phenotypes for the nuclear receptors in both, human and mouse do not only have influence on one metabolic pathway.
+For each nuclear receptor there were roughly about 8 to 10 phenotype annotations available, but only the ones which share the same function were written into the table.
+Nevertheless as an example the gene \textit{Nr2e1} without limitation regulates the expression for the same kind of molecule, a steroid hormone.
+This leads to the assumption that the nuclear receptors have somehow a similar influence on the individual in each species, albeit there are only few annotated by this time.
 ~~~~~~~\\
 ~~~~~~~\\
-%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%
-%which are the joined phenotypes, what's the consequence? e.g. what does it mean that mouse + human share steroid hormone? 
-%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%
 ~~~~~~~\\
@@
 Nevertheless, the mouse public databases are far less comprehensive than the UCSC, for instance, allowing for only a generalised comparison between the two species, relative to the phenotypic expression in case of a variation in one of the same nuclear receptors.
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-%what are the similarites, differences? How does the mouse study involve human phenotypes?
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%
+~~~~~~~\\
+~~~~~~~\\
+The mapping of phenotype information between mouse and human showed that regardless if the genes are thomolog, their function can differ between these two species, at least with the information we have by this time on the human nuclear receptors.
+As long as the human nuclear receptors are not enough researched, mapping the nuclear receptor variatons from mouse on the human can only result in an indicator to get information about what protein the gene could code for or in which biological process is it likely to be involved.
 

### 2015-03-15 article/article.tex
commit

@@
 	NZB/BlNJ  & Nr5a1 (nuclear receptor subfamily 5, group A, member 1) & succinylcarnitine relative abundance  &  3.8\\
-	C57L/J & Esr2 (estrogen receptor 2 beta) & 
-	relative size of perivascular immune cell clusters &  3.5\\
-	CE/J & Rorb (RAR-related orphan receptor beta) & 
-	tigloylglycine (16h fast), relative abundance&  3.3\\
-	NOD/\\ShiLtJ & Ppard (peroxisome proliferator activator receptor delta) & 
-	percentage of parasites in brain relative to all organs tested &  3.1\\ 
+	C57L/J & Esr2 (estrogen receptor 2 beta) & relative size of perivascular immune cell clusters &  3.5\\
+	CE/J & Rorb (RAR-related orphan receptor beta) & tigloylglycine (16h fast), relative abundance&  3.3\\
+	NOD/\\ShiLtJ & Ppard (peroxisome proliferator activator receptor delta) & percentage of parasites in brain relative to all organs tested &  3.1\\ 
 	BUB/BnJ & Esr1 (estrogen receptor 1 alpha) & 3-dehydrocarnitine  relative abundance &  2.7\\   
+	tdb & Esrrg ( ... ) & tbd &  tbd\\ 
 	\hline
@@
 Moreover, Figures~\ref{fig:mpi_pheotypes_strain_zscore2} and~\ref{fig:mpi_pheotypes_strain_zscore2_neg} provide an insight into the exact association of each phenotype with the corresponding nuclear receptor genes.
-Therefore, there are ...
-genes associated with the extreme high phenotypes and 5 genes associated with the extreme low phenotypes, respectively, as following: \textit{Rora}, \textit{Esr1}, \textit{Esrrg}, \textit{Thrb} and \textit{Nr3c2}.
+Therefore, there are 4 genes associated with the extreme high and low phenotypes respectively, as following: \textit{Rora - RAR-related orphan receptor alpha}, \textit{Esr1 - estrogen receptor 1 alpha}, \textit{Esrrg - estrogen-related receptor gamma}, \textit{Thrb - thyroid hormone receptor, beta} and a fifth gene associated only with extreme low phenotypes - \textit{Nr3c2}.
 \begin{figure}[H]

### 2015-03-23 article/article.tex
article

@@
 The goal of this analysis is to find NR-associated SNPs in mice that influence changes in biological parameters such as body weight, body fat and other phenotypic traits.
-Furthermore, these findings will be coupled to phenotypes observed in mice with a targeted or spontaneous mutation of the nuclear receptor and thus provide additional indication for a putative functionality of the investigated SNPs.}}
+Furthermore, these findings will be coupled to phenotypes observed in human metabolic pathways and thus provide an insight into the implications of the investigated SNPs in human disease and drug development processes.}}
 
@@
 ~~~~~~~\\  
-This paper makes use of the publicly available data in the Mouse Genome Informatics database and Mouse Phenome Database, respectively, in order to highlight changes in various biological parameters in mice under the influence of the NR-associated SNPs.
-Furthermore, these findings can be mapped to genotype - phenotype associations in humans, for studying the human biology and disease.
+This paper makes use of the publicly available data in the Mouse Genome Informatics database~\cite{mgi} and Mouse Phenome Database~\cite{mpd}, respectively, in order to highlight changes in various biological parameters in mice under the influence of the NR-associated SNPs.
+Furthermore this enables the study of the human biology and disease, by mapping these findings to genotype - phenotype associations in humans found in the HMDB (Human Metabolome Database)~\cite{hm1},~\cite{hm2}, EHMN (Edinburgh Human Metabolic Network)~\cite{hm3}, Recon X (Reconstruction of Human Metabolism)~\cite{hm4} and KEGG (Kyoto Encyclopedia of Genes and Genomes)~\cite{hm5},~\cite{hm6}.
 
@@
 Similarly, diverging mouse strains present non-identical phenotypes for the genetic parameters.
-
-\subsection{Mouse nuclear receptors}
-
+~~~~~~~\\
+~~~~~~~\\  
 The analysis presented in this paper was based upon a dataset consisting of the 49 nuclear receptors of mouse~\cite{proteomic} and their associated genes (242 unique gene Ids, 49 gene names / symbols); for a full list of gene names, see Appendix~\ref{an:appendix}).
@@
 More than 1330 strains were examined, providing annotations for over 3500 phenotype and 1.8 billion genotypes.
+The MPD database is more detailed and comprehensive than the MGI, so that the phenotypes found in MGI can be traced back to MPD.
+However, the MPD strictly associates individual phenotypes to their corresponding strain, posing difficulties in the mapping process between the mouse nuclear receptors and the MGI data.
+
+\subsection{Outcast: HMDB, EHMN, Recon X, KEGG}
+The nuclear receptor variation has been widely studied not only in the mouse, but also in the human genome.
+The research hereby refers to two published genome-wide association studies in this field of interest, namely \textit{Kora}\footnote{\url{http://epi.helmholtz-muenchen.de/kora-gen}, \today}  - a GWAS study initiated at the Helmholtz Zentrum in Munich involving approximately 18,000 adults from Southern Germany - and \textit{TwinsUK}\footnote{\url{http://www.twinsuk.ac.uk/}, \today}, being the biggest adult twin genetic registry in the UK.
 ~~~~~~~\\
 ~~~~~~~\\
-The MPD database is more detailed and comprehensive than the MGI, so that the phenotypes found in MGI can be traced back to MPD.
-However, the MPD strictly associates individual phenotypes to their corresponding strain, posing difficulties in the mapping process between the mouse nuclear receptors and the MGI data.
+Using the information provided by these studies regarding the correlations between different SNPs and metabolites to metabolic pathways, this research aimed to identify the similarities between the nuclear receptors and their associated phenotypes in human and in the mouse, respectively.
+Therefore, it makes use of the information about metabolite names, classes and associated diseases, pathways and tissues for metabolites found in the HMDB (Human Metabolome Database)~\cite{hm1},~\cite{hm2}, EHMN (Edinburgh Human Metabolic Network)~\cite{hm3}, Recon X (Reconstruction of Human Metabolism)~\cite{hm4} and KEGG (Kyoto Encyclopedia of Genes and Genomes)~\cite{hm5},~\cite{hm6}.
 
@@
 Since annotation from multiple public databases and internal data were used, there was a need to design and implement a database supporting the research.
-The database structure thereby presents 7 tables and aims to facilitate the processes of gathering, parsing and statistical analysis of the nuclear receptors data, as well as to allow for a clear mapping between the nuclear receptor genes and the phenotypes and SNPs associated with them.
-The main challenge was to establish a connection between the MGI, MPD and UCSC tables.
-Therefore, there are 2 tables containing information from the MGI, 3 tables for MPD, 1 USCS table and the core table:
+The database structure thereby presents 10 tables and aims to facilitate the processes of gathering, parsing and statistical analysis of the nuclear receptors data, as well as to allow for a clear mapping between the nuclear receptor genes and the phenotypes and SNPs associated with them.
+The main challenge was to establish a connection between the MGI, MPD and HMDB tables.
+Therefore, there are 2 tables containing information from the MGI, 4 tables for MPD, 3 HMDB tables and the core table:
 \begin{itemize}
@@
 \item \textbf{mpd\_phenotypes} - stores phenotypic information in association with MPD strains; more than one phenotype can be associated with a specific strain.
+\item \textbf{kora} - stores metabolic information from the Kora study.
+\item \textbf{twins} - stores metabolic information from the Twins study in the UK.
+\item \textbf{human\_metabolites} - stores metabolic information from the HMDB, EHMN, Recon X and KEGG.  
 \end{itemize} 
@@
 	~~~~~~~\\
-	Database scheme for the 49 nuclear receptors with information about MGI, MPI and UCSC.}}
+	Database scheme for the 49 nuclear receptors with information about MGI, MPI and HMDB.}}
 	\label{fig:database}
 \end{figure}
+
 %------------------------------------------------
@@
 \subsection{MGI Statistics}
-
 \begin{figure}[H]
@@
 Figure~\ref{fig:mgi_pheotypes_distribution} illustrates the most significant phenotypes and their occurrence frequency across the 242 genes corresponding to the nuclear receptors in the mouse.
-With a count of 196 matches across the dataset, the most prominent phenotype correlates with the homeostatic metabolic processes, such as temperature regulation and pH-balance.
-Moreover, Figure~\ref{fig:mgi_pheotypes_nr} provides an insight into the exact association of each phenotype with the corresponding nuclear receptor genes, such that homeostatis, for instance, is prominently found in the following genes: \textit{Esr1}, \textit{Pparg}, \textit{Thrb}, \textit{Thra}, \textit{Vdr} etc.
+With a count of 196 matches across the dataset, the most prominent phenotype correlates with the homeostatic metabolic processes.
+Moreover, Figure~\ref{fig:mgi_pheotypes_nr} provides an insight into the exact association of each phenotype with the corresponding nuclear receptor genes, such that homeostatis, for instance, is prominently associated with mutations in the following genes: \textit{Esr1 - estrogen receptor 1 alpha}, \textit{Pparg - peroxisome proliferator-activated receptor gamma}, \textit{Thrb - thyroid hormone receptor, beta}, \textit{Thra - thyroid hormone receptor, alpha}, \textit{Vdr - vitamin D receptor gene} etc.
 Other phenotypes describe body size and growth features and are representative for the \textit{Pparg} genes.
-
 \begin{figure}[H]
@@
 \end{figure}
-~~~~~~~\\
 
 \subsection{MPD Statistics}
+The MPDatabase contains annotations for extreme phenotypes associated with different mouse strains.
+The determining factor in \textit{MPD} for associating a variation with a phenotypes is the Z-score, which describes the standard deviation of the standard distribution function.
+The range of the Z-score which indicates a significant deviation is below -2 (extremely low) and above 2 (extremely high).
+Between the range of -2 and 2, the Z-score does not deviate significantly enough from the mean to indicate any major changes and is therefore not relevant for this association.
+
 \begin{figure}[H]
@@
 ~~~~~~~\\
-
 \begin{figure}[H]
@@
 ~~~~~~~\\
-~~~~~~~\\
-The MPDatabase contains annotations for extreme phenotypes associated with different mouse strains.
-An extreme phenotype is described by a Z-score above 2 (very high) or below -2 (very low), relative to other measurement means.
 Based on these specifications, Figure~\ref{fig:mpi_pheotypes_distribution_zscore2} illustrates the most significant phenotypes and their occurrence frequency across the 242 genes corresponding to the nuclear receptors in the mouse, with a Z-score $>$ 2 (extreme high phenotypes).
@@
 In this case, the most significant phenotypes include \textit{body weight} as well, but the focus lies on several blood phenotypes (e.g.
-lymphocyte differential, red blood cell count, hemoglobin).
+lymphocyte differential, red blood cell count, haemoglobin).
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\linewidth]{pics/mpi_phenotypes_nr_zscore2.pdf}
+	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
+ 	\caption{\footnotesize{\textbf{MPD extreme phenotypes, having Z-score $>$ 2}. 
 	~~~~~~~\\
+	Mouse Phenome Database extreme phenotype occurrence frequency among the 49 nuclear receptors in the mouse, having Z-score $>$ 2}}
+	\label{fig:mpi_pheotypes_strain_zscore2}
+\end{figure}
 ~~~~~~~\\
-Table~\ref{tab:top10pheno} shows the strains and gene names associated with the highest ranking \textit{MPD} phenotypes relative to their Z-scores.
-As seen in the table, the phenotypes are not directly associated with the expressed proteins, but rather with the observed abnormal condition in the mouse strains.
-%Furthermore the denoted phenotypes in the table of \textit{MPD} show nevertheless a more detailed description than \textit{MGI}.
+Moreover, Figures~\ref{fig:mpi_pheotypes_strain_zscore2} and~\ref{fig:mpi_pheotypes_strain_zscore2_neg} provide an insight into the exact association of each phenotype with mutation in the corresponding nuclear receptor genes.
+Therefore, there are 4 genes associated with the extreme high and low phenotypes respectively, as following: \textit{Rora - RAR-related orphan receptor alpha}, \textit{Esr1 - estrogen receptor 1 alpha}, \textit{Esrrg - estrogen-related receptor gamma}, \textit{Thrb - thyroid hormone receptor, beta} and a fifth gene associated only with extreme low phenotypes - \textit{Nr3c2}.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\linewidth]{pics/mpi_phenotypes_nr_zscore2_neg.pdf}
+	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
+ 	\caption{\footnotesize{\textbf{MPD extreme phenotypes, having Z-score $<$ -2}. 
+	~~~~~~~\\
+	Mouse Phenome Database extreme phenotype occurrence frequency among the 49 nuclear receptors in the mouse, having Z-score $<$ -2}}
+	\label{fig:mpi_pheotypes_strain_zscore2_neg}
+\end{figure}
+~~~~~~~\\
+The Z-score in \textit{MPD} derives from measurements of metabolites or observations in specific mouse strains.
+In this regard, Table~\ref{tab:top10pheno} shows a snippet of strains and their gene names associated based on their Z-score.
+This shows that the phenotypes are not directly associated with the expressed proteins, but rather with the observed abnormal condition in the mouse strains.
 \begin{table}[H]
@@
 	NOD/\\ShiLtJ & Ppard (peroxisome proliferator activator receptor delta) & percentage of parasites in brain relative to all organs tested &  3.1\\ 
+	AKR/J & Esrrg (estrogen-related receptor gamma) & allantoin (16h fast), relative abundance &  2.8\\
 	BUB/BnJ & Esr1 (estrogen receptor 1 alpha) & 3-dehydrocarnitine  relative abundance &  2.7\\    
-	tdb & Esrrg ( ... ) & tbd &  tbd\\ 
 	\hline
 \end{tabulary}
-\caption{\footnotesize{\textbf{Top 10 extreme phenotypes.} 
+\caption{\footnotesize{\textbf{A snippet of extreme phenotypes.} 
 ~~~~~~~\\
@@
 \end{table}
-~~~~~~~\\
-Moreover, Figures~\ref{fig:mpi_pheotypes_strain_zscore2} and~\ref{fig:mpi_pheotypes_strain_zscore2_neg} provide an insight into the exact association of each phenotype with the corresponding nuclear receptor genes.
-Therefore, there are 4 genes associated with the extreme high and low phenotypes respectively, as following: \textit{Rora - RAR-related orphan receptor alpha}, \textit{Esr1 - estrogen receptor 1 alpha}, \textit{Esrrg - estrogen-related receptor gamma}, \textit{Thrb - thyroid hormone receptor, beta} and a fifth gene associated only with extreme low phenotypes - \textit{Nr3c2}.
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\linewidth]{pics/mpi_phenotypes_nr_zscore2.pdf}
-	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
- 	\caption{\footnotesize{\textbf{MPD extreme phenotypes, having Z-score $>$ 2}. 
-	~~~~~~~\\
-	Mouse Phenome Database extreme phenotype occurrence frequency among the 49 nuclear receptors in the mouse, having Z-score $>$ 2}}
-	\label{fig:mpi_pheotypes_strain_zscore2}
-\end{figure}
-~~~~~~~\\
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\linewidth]{pics/mpi_phenotypes_nr_zscore2_neg.pdf}
-	\captionsetup{margin=12pt,format=plain,font=footnotesize,labelfont=bf}
- 	\caption{\footnotesize{\textbf{MPD extreme phenotypes, having Z-score $<$ -2}. 
-	~~~~~~~\\
-	Mouse Phenome Database extreme phenotype occurrence frequency among the 49 nuclear receptors in the mouse, having Z-score $<$ -2}}
-	\label{fig:mpi_pheotypes_strain_zscore2_neg}
-\end{figure}
 
 %------------------------------------------------
+
 \subsection{Nuclear receptors in the human genome}
-The nuclear receptor variation has been widely studied not only in the mouse, but also in the human genome.
-The research hereby refers to two published genome-wide association studies in this field of interest, namely \textit{Kora}\footnote{\url{http://epi.helmholtz-muenchen.de/kora-gen}, \today}  - a GWAS study initiated at the Helmholtz Zentrum in Munich involving approximately 18,000 adults from Southern Germany - and \textit{TwinsUK}\footnote{\url{http://www.twinsuk.ac.uk/}, \today}, being the biggest adult twin genetic registry in the UK.
+This research shows that there are only a few phenotypes shared between the mouse and human.
+The reason therefore is that the phenotypes for the nuclear receptors in both human and mouse are involved in more than one metabolic pathway.
+For each nuclear receptor there were roughly 8 to 10 phenotypic annotations available, but only the ones which share the same function were mentioned in Tables~\ref{tab:compHuMoMe} and~\ref{tab:compHuMo}.
 ~~~~~~~\\
 ~~~~~~~\\
-Using the information provided by these studies regarding the correlations between different SNPs and metabolites to metabolic pathways, this research aimed to identify the similarities between the nuclear receptors and their associated phenotypes in human and in the mouse, respectively.
-Due to missing or incomplete annotations for the nuclear receptors in the mouse, comparisons could only be made between the phenotypes extracted from the MGI and MPD databases (see tables \textit{mgi\_phenotypes} and \textit{mpg\_phenotypes} from the internal database~\ref{fig:database}).
-Table \ref{tab:compHuMo} shows a list of nuclear receptors found in both the mouse and human and which present a similar phenotype or share the same biological function.
-As seen there are only a few phenotypes shared between these two species.
-The reason therefore is that the phenotypes for the nuclear receptors in both, human and mouse do not only have influence on one metabolic pathway.
-For each nuclear receptor there were roughly about 8 to 10 phenotype annotations available, but only the ones which share the same function were written into the table.
-Nevertheless as an example the gene \textit{Nr2e1} without limitation regulates the expression for the same kind of molecule, a steroid hormone.
-This leads to the assumption that the nuclear receptors have somehow a similar influence on the individual in each species, albeit there are only few annotated by this time.
-~~~~~~~\\
-~~~~~~~\\
-~~~~~~~\\
+Due to issues concerning the mapping of the mouse and human phenotypes relative to nuclear receptors, only two genes were found to have the same biological function in the mouse and in human - see Table~\ref{tab:compHuMoMe}.
+On the other hand, Table~\ref{tab:compHuMo} shows a list of nuclear receptors from publications found in both the mouse and human and which present a similar phenotype or share the same biological function.
 \begin{table}[H]
@@
 	Essrg  & carnitine, \textit{lysine metabolism} & glutaroyl carnitine, \textit{lysine metabolism}\\
+	\hline
+\end{tabulary}
+\caption{\footnotesize{\textbf{Nuclear receptors in human and in the mouse} 
+~~~~~~~\\
+	Nuclear receptors in human and in the mouse which share a similar phenotype or biological function.}} 
+\label{tab:compHuMoMe}
+\end{table}
+\begin{table}[H]
+\begin{tabulary}{\linewidth}{ LLL }
+\hline
+	\textbf{Nuclear Receptor Gene} & \textbf{Mouse phenotype}  & \textbf{Human phenotype}
+	\\  
+	\hline \hline 
 	Thrb  & \textit{thyroid hormone resistance} & \textit{thyroid hormone resistance}\\
@@
 \end{tabulary}
-\caption{\footnotesize{\textbf{Nuclear receptors with their coding protein or phenotype} 
+\caption{\footnotesize{\textbf{Nuclear receptors in human and in the mouse} 
 ~~~~~~~\\
@@
 As a result, it has been proved that nuclear receptors have a great influence over the extreme phenotypes showcased in the mouse, as well as in human.
+
+\subsection{MGI - MPD correlation}
+
+In spite of the ascending progress in this field of research, there annotation data provided in the mouse public databases - \textit{Mouse Genome Informatics} and \textit{Mouse Phenome Database} on nuclear receptors is still rather limited.
+However, the two databases provide complementary information: the phenotypes annotated in \textit{MGI} describe roughly the biological systems in which they occur rather than the description of differences between non-variants, as of \textit{MPD}.
+On the other hand, the nuclear receptor database for the human genome is far more detailed, comprising also information regarding the metabolic pathways associated with various phenotypes.
 ~~~~~~~\\
 ~~~~~~~\\
-In spite of the ascending progress in this field of research, there annotation data provided in the mouse public databases - \textit{Mouse Genome Informatics} and \textit{Mouse Phenome Database} on nuclear receptors is still rather limited.
-However, the two databases provide complementary information: the phenotypes annotated in \textit{MGI} describe roughly the biological systems in which they occur rather than the description of differences between non-variants, as of \textit{MPD}.
-On the other hand, the UCSC database for the human genome is far more detailed, comprising also information regarding the metabolic pathways associated with various phenotypes.
+According to the MGI database, as shown in Figures~\ref{fig:mgi_pheotypes_distribution} and~\ref{fig:mgi_pheotypes_nr}, the most prominent phenotype correlates with the homeostatic metabolic processes, such as temperature regulation, pH-balance as well as lipid and glucose homeostasis and is associated with mutations in the following genes: \textit{Esr1 - estrogen receptor 1 alpha}, \textit{Pparg - peroxisome proliferator-activated receptor gamma}, \textit{Thrb - thyroid hormone receptor, beta}, \textit{Thra - thyroid hormone receptor, alpha}, \textit{Vdr - vitamin D receptor gene}.
+The Esr1 - Estrogen receptor alpha gene is known to mediates effects of estrogen on glucose homeostasis~\cite{esr1_discussion}.
+Also the thyroid hormone, acting through thyroid hormone receptors has been proven to be a key regulator of metabolic homeostasis~\cite{thra_discussion}.
 ~~~~~~~\\
 ~~~~~~~\\
+Comparably, the MPD database presents rather different phenotypes, associated with an extreme Z-score.
+Therefore, Figures~\ref{fig:mpi_pheotypes_distribution_zscore2} and Figure~\ref{fig:mpi_pheotypes_distribution_zscore2_neg} illustrate extreme high and low phenotypes, respectively.
+Phenotypes presented with an extreme high effect include  \textit{body weight}, \textit{kidney weight} and \textit{cholesterol}, while those characterised by an extreme low effect on the organism depict blood phenotypes (e.g.
+\textit{lymphocyte differential}, \textit{red blood cell count}, \textit{haemoglobin}).
+Nevertheless, there are 4 genes associated with the extreme high and low phenotypes respectively, as following: \textit{Rora - RAR-related orphan receptor alpha}, \textit{Esr1 - estrogen receptor 1 alpha}, \textit{Esrrg - estrogen-related receptor gamma}, \textit{Thrb - thyroid hormone receptor, beta} and a fifth gene associated only with extreme low phenotypes - \textit{Nr3c2}, as supported by a recent review~\cite{blood_discussion}.
+Studies~\cite{esr12_discussion},~\cite{esrrg_discussion} have shown that the oestrogen receptor genes Esr1 and Esrrg are indeed involved in body growth processes, by playing an essential role in cell division and DNA recombination.
+
+\subsection{Mouse - Human correlation}
+
 This research presents the most significant phenotypes found in the mouse, based on the 49 nuclear receptor variation, as well as a mapping of these phenotypes on the human phenotypic expression, providing insight into possible human disease conditions.
-Nevertheless, the mouse public databases are far less comprehensive than the UCSC, for instance, allowing for only a generalised comparison between the two species, relative to the phenotypic expression in case of a variation in one of the same nuclear receptors.
+Nevertheless, the mouse public databases are far less comprehensive than the HMDB, for instance, allowing for only a generalised comparison between the two species, relative to the phenotypic expression in case of a variation in one of the same nuclear receptors.
 ~~~~~~~\\
 ~~~~~~~\\
-The mapping of phenotype information between mouse and human showed that regardless if the genes are thomolog, their function can differ between these two species, at least with the information we have by this time on the human nuclear receptors.
-As long as the human nuclear receptors are not enough researched, mapping the nuclear receptor variatons from mouse on the human can only result in an indicator to get information about what protein the gene could code for or in which biological process is it likely to be involved.
+Tables~\ref{tab:compHuMoMe} and~\ref{tab:compHuMo} present genes which were found to have the same biological function in the mouse and in human.
+On the one hand, gene \textit{Nr2e1} regulates the expression for the same kind of molecule, a steroid hormone; this is consistent to the conclusions of several published studies\footnote{\url{http://www.phosphosite.org/proteinAction.do?id=2204525&showAllSites=true}, \today}, \footnote{\url{http://www.phosphosite.org/proteinAction.do?id=2204526&showAllSites=true}, \today}  and leads to the assumption that the nuclear receptors have somehow a similar influence on the individual in each species, although there are only few annotated by this time.
+On the other hand, the overly researched gene Essrg has also been found,~\cite{esrrg2_discussion} to play an important role in the lysine metabolism in both human and the mouse.
+~~~~~~~\\
+~~~~~~~\\
+The mapping of phenotype information between mouse and human showed that regardless if the genes are homolog, their function can differ between these two species, at least with the information we have by this time on the human nuclear receptors.
+As long as the human nuclear receptors are not enough researched, mapping the nuclear receptor variations from mouse on the human can only result in an indicator to get information about what protein the gene could code for or in which biological process is it likely to be involved.
 
 %------------------------------------------------
+
 \section{Acknowledgments} % The \section*{} command stops section numbering
